### 2024-04-09 data-prep/04-Creating-avail_2011-16.Rmd

---
title: "Testing-doc-2016run"
output: html_document
date: "2024-02-16"
---
# 2011 calcs

```{r load-packages, cache=FALSE}
library(dplyr)
library(sf)
library(ggplot2)
library(tidyverse)
library(units)
library(patchwork)
library(ggspatial)
options(scipen = 999)
```

```{r sp_avail-function}
#defining the spatial availability function, this prints all the intermediates
sp_avail_detailed <- function(x, o_id, d_id, pop, opp, pop_type, r, f, alpha = 1){

  o_id <- rlang::enquo(o_id)
  d_id <- rlang::enquo(d_id)
  pop <- rlang::enquo(pop)
  opp <- rlang::enquo(opp)
  pop_type <- rlang::enquo(pop_type)
  r <- rlang::enquo(r)
  f <- rlang::enquo(f)

  sum_pop <- x %>%
    dplyr::distinct(!!o_id, !!pop_type,
                    .keep_all = TRUE) %>%
    dplyr::mutate(sum_pop = !!r*(!!pop)^alpha) %>%
    dplyr::pull(sum_pop) %>%
    sum()

  f_p <- dplyr::pull(x, !!r) * dplyr::pull(x, !!pop)^alpha / sum_pop

  sum_impedance <- x %>%
    dplyr::group_by(!!d_id) %>%
    dplyr::summarize(sum_impedance = sum(!!f))

  x <- x %>%
    dplyr::left_join(sum_impedance,
                     by = rlang::as_name(d_id))

  f_c <- dplyr::pull(x, !!f) / x$sum_impedance

  x$f_c <- f_c
  x$f_p <- f_p

  sum_pa <- x %>%
    dplyr::group_by(!!d_id) %>%
    dplyr::summarize(sum_pa= sum(f_p * f_c))

  x <- x %>%
    dplyr::left_join(sum_pa,
                     by = rlang::as_name(d_id))
  f_t <- (f_p * f_c) / dplyr::pull(x, sum_pa)

  x %>%
    dplyr::mutate(V_ij = !!opp * !!f_t)
}
```

```{r}
load(file="data-products/cen_data_2011_hamiltononly.Rdata")
cen_data_2011_hamiltononly <- cen_data_2011_1 |> filter(Region.Name == "Hamilton") #NOTE: GEOUID 35250242, 35250133, 35250457 do not have any child population
rm("cen_data_2011_1")
load(file="data-products/cen_data_2016_hamiltononly.Rdata")
cen_data_2016_hamiltononly <- cen_data_2016_1 |> filter(Region.Name == "Hamilton") #NOTE: GEOUID 35250133, 35250398, 35250427 do not have any child population
rm("cen_data_2016_1")

ham_city_bound <- st_read("data-inputs/Boundaries/City_Boundary.shp") |> st_transform(st_crs(cen_data_2011_hamiltononly))
ham_community_bounds <- st_read("data-inputs/Boundaries/Community_Boundaries.shp") |> st_transform(st_crs(cen_data_2011_hamiltononly))
ggh_taz <- st_read("data-inputs/TTS/tts06_83_region.shp") |> st_transform(st_crs(cen_data_2011_hamiltononly))
```

```{r}
cen_data_2011_hamiltononly <- 
  cen_data_2011_hamiltononly %>% dplyr::select( #removing the variables we don't need for this analysis
  -c("Shape.Area", "Quality.Flags", "Type", "CD_UID", "NHS.Non.Return.Rate", "CSD_UID", "CT_UID", "CT_UID","NHS.Non.Return.Rate.1", "v_CA11N_2564..Median.after.tax.household.income..", "v_CA11F_17..15.to.19.years")) %>%
  rename( #renaming the variables
    "DA_households" = "Households",
    "DA_dwellings" = "Dwellings",
    "DA_tot_pop" = "Population",
    "DA_med_AT_family_income" = "v_CA11N_2458..Median.after.tax.family.income..",
    "DA_med_AT_lone-family_income" = "v_CA11N_2476..Median.after.tax.family.income..",
    "DA_prev_LIMAT_under18" = "v_CA11N_2609..Less.than.18.years..",
    "DA_pop_5to9" = "v_CA11F_11..5.to.9.years",
    "DA_pop_10to14" = "v_CA11F_14..10.to.14.years") %>%
  mutate(DA_pop_5to14 = DA_pop_5to9+(DA_pop_10to14)) |> #*0.9)) %>%
  dplyr::select(-c("DA_pop_5to9","DA_pop_10to14"))

cen_data_2016_hamiltononly <- 
  cen_data_2016_hamiltononly %>% dplyr::select( #removing the variables we don't need for this analysis
  -c("Shape.Area",  "Type", "CD_UID", "CSD_UID", "CT_UID", "CT_UID", "v_CA16_2398..Median.after.tax.income.of.households.in.2015....", "v_CA16_64..15.to.19.years")) %>%
  rename( #renaming the variables
    "DA_households" = "Households",
    "DA_dwellings" = "Dwellings",
    "DA_tot_pop" = "Population",
    "DA_med_AT_family_income" = "v_CA16_2448..Median.after.tax.income.of.economic.families.in.2015....",
         "DA_med_AT_lone-family_income" = "v_CA16_2460..Median.after.tax.income.of.lone.parent.economic.families.in.2015....",
         "DA_prev_LIMAT_under18" = "v_CA16_2543..0.to.17.years....",
         "DA_pop_5to9" = "v_CA16_25..5.to.9.years",
         "DA_pop_10to14" = "v_CA16_43..10.to.14.years") %>% mutate(DA_pop_5to14 = DA_pop_5to9+(DA_pop_10to14)) |> #*0.9)) %>%
         dplyr::select(-c("DA_pop_5to9","DA_pop_10to14"))
```

```{r}
library(hamONdests)
data(Schools_2011_2016)
ALL_SCHOOLS_Elem <- Schools_2011_2016 %>% filter(Level == "Elementary")
rm("Schools_2011_2016")
#data(School_Catchments_2011_2016)
load(file="data-products/TTS_gamma2011_motor.Rdata")
load(file="data-products/TTS_gamma2016_motor.Rdata")
load(file="data-products/TTS_exp2011_nonmotor.Rdata")
load(file="data-products/TTS_exp2016_nonmotor.Rdata")
```

```{r}
#add GTA06 IDs to schools - "GTA06_Dest"}
cols_TAZIDs <- sp::over(as_Spatial(ALL_SCHOOLS_Elem |> st_transform(st_crs(ggh_taz))),
                              as_Spatial(ggh_taz |> dplyr::select(GTA06) |> rename("GTA06_Dest" = "GTA06")))

ALL_SCHOOLS_Elem <- cbind(ALL_SCHOOLS_Elem,cols_TAZIDs)
rm(cols_TAZIDs)
```

```{r}
load(file="data-products/OD_tt_car_2011_1.Rdata")
load(file="data-products/OD_tt_car_2011_2.Rdata")
load(file="data-products/OD_tt_car_2011_3.Rdata")
load(file="data-products/OD_tt_car_2011_4.Rdata")

load(file="data-products/OD_tt_walk_2011_1.Rdata")
load(file="data-products/OD_tt_walk_2011_2.Rdata")
load(file="data-products/OD_tt_walk_2011_3.Rdata")
load(file="data-products/OD_tt_walk_2011_4.Rdata")

load(file="data-inputs/TTS/od_trips_perc_2011.Rdata")

#renaming travel time matrices 
ttm_car_2011 <- bind_rows(output_OD_car1, output_OD_car2, output_OD_car3, output_OD_car4) %>% mutate(mode = "c")
rm(output_OD_car1, output_OD_car2, output_OD_car3, output_OD_car4)

ttm_walk_2011 <- bind_rows(output_OD_walk_elem1,output_OD_walk_elem2,output_OD_walk_elem3,output_OD_walk_elem4) %>% mutate(mode = "w")
rm(output_OD_walk_elem1,output_OD_walk_elem2,output_OD_walk_elem3,output_OD_walk_elem4)
```

```{r}
load(file="data-inputs/Parcels/RED_LU_centroids_2011.RData") #parcels

#drop all parcel points that do not have a GeoUID (i.e., a DA identifier -- this means they fall outside of Hamilton). It's 10 for 2011 and 11 for 2016
RED_LU_centroids_2011 <- RED_LU_centroids_2011 %>% drop_na(GeoUID) %>%
  group_by(ID) %>%
  mutate(res_units = sum(SINGLE_FAM,SEMI_DET,ROW_THOUSE,APARTMENT)) %>%
  dplyr::select(-c("SINGLE_FAM","SEMI_DET","ROW_THOUSE","APARTMENT",
                   "TERANET", "LOCATION", "LUC1", "LUC2", "TYPE", "ROLL_NUM"))
```

```{r}
#add census data to the parcel centroids:
RED_LU_centroids_2011 <- RED_LU_centroids_2011 %>%
  merge(cen_data_2011_hamiltononly %>% st_drop_geometry(), by="GeoUID") 
```

```{r}
#create OD matrix -- each parcel to each school, and then drop all OD pairs that don't have a travel time ()
OD_2011 <- RED_LU_centroids_2011 %>% 
  cross_join(ALL_SCHOOLS_Elem %>% st_drop_geometry() |> filter(Year != 2016) %>% dplyr::select(c("SchoolID", "OTGC2011", "GTA06_Dest")))

#add travel time, per mode
test <- OD_2011 %>% mutate(ID = as.character(ID),
                           SchoolID = as.character(SchoolID)) %>%
  left_join(ttm_car_2011, by=c("ID" = "from_id",
                               "SchoolID" = "to_id")) %>% 
  drop_na(travel_time_p50)

OD_2011 <- OD_2011 %>% mutate(ID = as.character(ID),
                           SchoolID = as.character(SchoolID)) %>%
  left_join(ttm_walk_2011 |> filter(travel_time_p50 <= 27), by=c("ID" = "from_id",
                                "SchoolID" = "to_id")) %>% 
  drop_na(travel_time_p50)

OD_2011 <- rbind(test,OD_2011) 
rm(test, ttm_car_2011, ttm_walk_2011)
```
```{r}
#reduce dimensions
OD_2011 <- OD_2011 %>% dplyr::select(-c("LUC1_DESC", "LUC2_DESC", "res_units", "DA_dwellings", "CMA_UID", "Region.Name", "geometry","PROP_AREA", "Area..sq.km."))

# Check the mean and median travel times by mode, the first is at the level of zn1259 and the second is at zn208
OD_2011 %>% 
  group_by(mode) %>% 
  summarize(mean_tt = mean(travel_time_p50),
            median_tt = median(travel_time_p50),
            min_tt = min(travel_time_p50),
            max_tt = max(travel_time_p50))
```

```{r}
#now let's add the communtiy names to the TAZ and summarize the % of students who use what mode by community. 
ggh_taz_communities <- st_intersection(st_centroid(ham_taz) |> st_transform(st_crs(ggh_taz)), 
                                       ham_community_bounds |> st_transform(st_crs(ggh_taz))) |>
  dplyr::select(c("GTA06", "REGION", "COMMUNITY_"))

od_trips_perc_2011 <- od_trips_perc_2011 %>% left_join(ggh_taz_communities |> st_drop_geometry() |> mutate(GTA06 = as.character(GTA06)), by=c("Origin"="GTA06"))

od_trips_perc_comm_2011 <- od_trips_perc_2011 %>% group_by(COMMUNITY_) %>%
  summarize(TAZOrig_car_trips = sum(TAZOrig_car_trips),
            TAZOrig_walk_trips = sum(TAZOrig_walk_trips)) %>% 
  mutate(TAZOrig_car_perc = TAZOrig_car_trips/(TAZOrig_car_trips + TAZOrig_walk_trips),
         COMMUNITY = case_when(COMMUNITY_ == "Ancaster" ~ "AN",
                                                         COMMUNITY_ == "Dundas" ~ "DU",
                                                         COMMUNITY_ == "Flamborough" ~ "FL",
                                                         COMMUNITY_ == "Glanbrook" ~ "GL",
                                                         COMMUNITY_ == "Hamilton" ~ "HA",
                                                         COMMUNITY_ == "Stoney Creek" ~ "SC")) |>
  dplyr::select(-c("COMMUNITY_"))

od_trips_perc_comm_2011
```

```{r}
# Lets add the % of car-mode used to the full OD list and then start transforming the frame for Spatial Availability calc.
OD_2011 <- OD_2011 |> left_join(od_trips_perc_comm_2011 |> dplyr::select(c("TAZOrig_car_perc", "COMMUNITY")), 
                     by = "COMMUNITY") %>%
  mutate(
  DA_pop_5to14 = ifelse(is.na(DA_pop_5to14), 0, DA_pop_5to14),
  DA_households = ifelse(is.na(DA_households), 0, DA_households))

#'mode_count' identifies if a OD pair (ID to SchoolID) is either just car (1) or car and walk (2). Join this to the full OD list.
OD_2011_mode_count <- OD_2011 |>
  group_by(ID, SchoolID) %>% 
  summarise(mode_count = n())

OD_2011 <- OD_2011 %>% left_join(OD_2011_mode_count, by = c("ID", "SchoolID"))
rm(OD_2011_mode_count)
```

```{r}
#now, here we make the assumption on how many students live in each parcel. So since we assume that any parcel could potentially go to any school - I think it makes sense to also assume that the most granular level of population data we have (DA-level) is equally assigned to each parcel and then equally assigned to each OD flow based on it's mode-split (on average ~80/20 car/walk).

#Count the number of ID's for each OD pair -- this can be seen as the number of 'trips' being made from a parcel. There are 147 elementary schools in 2011 -- all parcels can reach all schools by car in 60 mins. So if a parcel only has 1 mode_count (car) it will have a "DA_ID_count" of 147 (the number of OD home-to-all-school trips). This is the number that the rate of Students in DA per OD flows in the DA ("IDs_in_a_DA_count") is multipled by.
DA_ID_count <- OD_2011 |> group_by(ID) |> summarise(DA_ID_count = n()) 
OD_2011 <- OD_2011 |> left_join(DA_ID_count, by="ID")

#Here we calculate the number of OD trips per O (ID) that occur in each DA. This is how much the # of students in a DA is divded by for each OD. 
IDs_in_a_DA_count <- OD_2011 |> group_by(GeoUID) |> summarise(IDs_in_a_DA_count = sum(DA_ID_count)) 
OD_2011 <- OD_2011 |> left_join(IDs_in_a_DA_count, by="GeoUID")

#We calculate the _kid-pop_ per parcel and then we adjusted it based on if the OD flow is 'walking' or 'car'
OD_2011 <- OD_2011 |> mutate(kid_pop = DA_pop_5to14/IDs_in_a_DA_count*DA_ID_count, 
               kid_pop = ifelse(is.na(kid_pop) | is.nan(kid_pop) | kid_pop == Inf, 0, kid_pop),
               kid_popadj = case_when(mode == "w" & mode_count == 2 ~ (1.0000000-TAZOrig_car_perc)*2*kid_pop,
                                         mode == "c" & mode_count == 2 ~ (TAZOrig_car_perc)*2*kid_pop,
                                         mode == "c" & mode_count == 1 ~ kid_pop),
                  kid_popadj = ifelse(is.na(kid_popadj) | is.nan(kid_popadj) | kid_popadj == Inf, 0, kid_popadj))
```

```{r}
#checks!, these should be the total number of kids (^the code above) -- 58265.
OD_2011$kid_pop %>% sum()
OD_2011$kid_popadj %>% sum()
OD_2011 |> group_by(GeoUID) |> summarize(DA_pop_5to14 = mean(DA_pop_5to14)) |> dplyr::select(DA_pop_5to14) |> sum()
```

```{r}
#now add the impedance function, etc to the full OD 
OD_2011 <- OD_2011 |> mutate(
  travel_time_p50 = ifelse(travel_time_p50 == 0, 0.5, travel_time_p50),
  r = 1,
  f = ifelse(mode == "c", stats::dgamma(
    travel_time_p50, gamma2011_motor$estimate["shape"], gamma2011_motor$estimate["rate"]), 
    stats::dexp(
      travel_time_p50, exp2011_nonmotor$estimate["rate"])))
```

```{r}
#calculate spatial availability. 
avail_2011 <- sp_avail_detailed(OD_2011, ID, SchoolID, kid_popadj, OTGC2011, mode, r, f, alpha = 1)
save(avail_2011, file="data-products/avail_2011.RData")
save(OD_2011, file="data-products/OD_2011.RData")
rm(OD_2011)
```

# 2016 calcs
```{r}
rm(list=ls())
```

```{r load-packages, include=FALSE, cache=FALSE}
library(dplyr)
library(sf)
library(ggplot2)
library(tidyverse)
library(units)
library(patchwork)
library(ggspatial)
options(scipen = 999)
```

```{r sp_avail-function,include=FALSE}
#defining the spatial availability function, this prints all the intermediates
sp_avail_detailed <- function(x, o_id, d_id, pop, opp, pop_type, r, f, alpha = 1){

  o_id <- rlang::enquo(o_id)
  d_id <- rlang::enquo(d_id)
  pop <- rlang::enquo(pop)
  opp <- rlang::enquo(opp)
  pop_type <- rlang::enquo(pop_type)
  r <- rlang::enquo(r)
  f <- rlang::enquo(f)

  sum_pop <- x %>%
    dplyr::distinct(!!o_id, !!pop_type,
                    .keep_all = TRUE) %>%
    dplyr::mutate(sum_pop = !!r*(!!pop)^alpha) %>%
    dplyr::pull(sum_pop) %>%
    sum()

  f_p <- dplyr::pull(x, !!r) * dplyr::pull(x, !!pop)^alpha / sum_pop

  sum_impedance <- x %>%
    dplyr::group_by(!!d_id) %>%
    dplyr::summarize(sum_impedance = sum(!!f))

  x <- x %>%
    dplyr::left_join(sum_impedance,
                     by = rlang::as_name(d_id))

  f_c <- dplyr::pull(x, !!f) / x$sum_impedance

  x$f_c <- f_c
  x$f_p <- f_p

  sum_pa <- x %>%
    dplyr::group_by(!!d_id) %>%
    dplyr::summarize(sum_pa= sum(f_p * f_c))

  x <- x %>%
    dplyr::left_join(sum_pa,
                     by = rlang::as_name(d_id))
  f_t <- (f_p * f_c) / dplyr::pull(x, sum_pa)

  x %>%
    dplyr::mutate(V_ij = !!opp * !!f_t)
}
```

```{r}
load(file="data-products/cen_data_2016_hamiltononly.Rdata")
cen_data_2016_hamiltononly <- cen_data_2016_1 |> filter(Region.Name == "Hamilton") #NOTE: GEOUID 35250133, 35250398, 35250427 do not have any child population
rm("cen_data_2016_1")

ham_city_bound <- st_read("data-inputs/Boundaries/City_Boundary.shp") |> st_transform(st_crs(cen_data_2016_hamiltononly))
ham_community_bounds <- st_read("data-inputs/Boundaries/Community_Boundaries.shp") |> st_transform(st_crs(cen_data_2016_hamiltononly))
ggh_taz <- st_read("data-inputs/TTS/tts06_83_region.shp") 
# |> st_transform(st_crs(cen_data_2011_hamiltononly))
```

```{r}
cen_data_2016_hamiltononly <- 
  cen_data_2016_hamiltononly %>% dplyr::select( #removing the variables we don't need for this analysis
  -c("Shape.Area",  "Type", "CD_UID", "CSD_UID", "CT_UID", "CT_UID", "v_CA16_2398..Median.after.tax.income.of.households.in.2015....", "v_CA16_64..15.to.19.years")) %>%
  rename( #renaming the variables
    "DA_households" = "Households",
    "DA_dwellings" = "Dwellings",
    "DA_tot_pop" = "Population",
    "DA_med_AT_family_income" = "v_CA16_2448..Median.after.tax.income.of.economic.families.in.2015....",
         "DA_med_AT_lone-family_income" = "v_CA16_2460..Median.after.tax.income.of.lone.parent.economic.families.in.2015....",
         "DA_prev_LIMAT_under18" = "v_CA16_2543..0.to.17.years....",
         "DA_pop_5to9" = "v_CA16_25..5.to.9.years",
         "DA_pop_10to14" = "v_CA16_43..10.to.14.years") %>% mutate(DA_pop_5to14 = DA_pop_5to9+(DA_pop_10to14)) |> #*0.9)) %>%
         dplyr::select(-c("DA_pop_5to9","DA_pop_10to14"))
```

```{r, include=FALSE}
library(hamONdests)
data(Schools_2011_2016)
ALL_SCHOOLS_Elem <- Schools_2011_2016 %>% filter(Level == "Elementary")
rm("Schools_2011_2016")
#data(School_Catchments_2011_2016)
load(file="data-products/TTS_gamma2016_motor.Rdata")
load(file="data-products/TTS_exp2016_nonmotor.Rdata")
```

```{r, include=FALSE}
#add GTA06 IDs to schools - "GTA06_Dest"}
cols_TAZIDs <- sp::over(as_Spatial(ALL_SCHOOLS_Elem |> st_transform(st_crs(ggh_taz))),
                              as_Spatial(ggh_taz |> dplyr::select(GTA06) |> rename("GTA06_Dest" = "GTA06")))
ALL_SCHOOLS_Elem <- cbind(ALL_SCHOOLS_Elem,cols_TAZIDs)
rm(cols_TAZIDs)
```

```{r}
load(file="data-products/OD_tt_car_2016_1.Rdata")
load(file="data-products/OD_tt_car_2016_2.Rdata")
load(file="data-products/OD_tt_car_2016_3.Rdata")
load(file="data-products/OD_tt_car_2016_4.Rdata")

load(file="data-products/OD_tt_walk_2016_1.Rdata")
load(file="data-products/OD_tt_walk_2016_2.Rdata")
load(file="data-products/OD_tt_walk_2016_3.Rdata")
load(file="data-products/OD_tt_walk_2016_4.Rdata")

load(file="data-inputs/TTS/od_trips_perc_2016.Rdata")

#renaming travel time matrices 
ttm_car_2016 <- bind_rows(output_OD_car1, output_OD_car2, output_OD_car3, output_OD_car4) %>% mutate(mode = "c")
rm(output_OD_car1, output_OD_car2, output_OD_car3, output_OD_car4)

ttm_walk_2016 <- bind_rows(output_OD_walk_elem1,output_OD_walk_elem2,output_OD_walk_elem3,output_OD_walk_elem4) %>% mutate(mode = "w")
rm(output_OD_walk_elem1,output_OD_walk_elem2,output_OD_walk_elem3,output_OD_walk_elem4)
```

```{r}
load(file="data-inputs/Parcels/RED_LU_centroids_2016.RData") #parcels

#drop all parcel points that do not have a GeoUID (i.e., a DA identifier -- this means they fall outside of Hamilton). It's 10 for 2011 and 11 for 2016
RED_LU_centroids_2016 <- RED_LU_centroids_2016 %>% drop_na(GeoUID) %>%
  group_by(ID) %>%
  mutate(res_units = sum(SINGLE_FAM,SEMI_DET,ROW_THOUSE,APARTMENT)) %>%
  dplyr::select(-c("SINGLE_FAM","SEMI_DET","ROW_THOUSE","APARTMENT",
                   "TERANET", "LOCATION", "LUC1", "LUC2", "TYPE", "ROLL_NUM"))
```

```{r}
#add census data to the parcel centroids:
RED_LU_centroids_2016 <- RED_LU_centroids_2016 %>%
  merge(cen_data_2016_hamiltononly %>% st_drop_geometry(), by="GeoUID") 
```

```{r}
#create OD matrix -- each parcel to each school, and then drop all OD pairs that don't have a travel time ()
OD_2016 <- RED_LU_centroids_2016 %>% 
  cross_join(ALL_SCHOOLS_Elem %>% st_drop_geometry() |> filter(Year != 2011) %>% dplyr::select(c("SchoolID", "OTGC2016", "GTA06_Dest")))

#add travel time, per mode
test <- OD_2016 %>% mutate(ID = as.character(ID),
                           SchoolID = as.character(SchoolID)) %>%
  left_join(ttm_car_2016, by=c("ID" = "from_id",
                               "SchoolID" = "to_id")) %>% 
  drop_na(travel_time_p50)

OD_2016 <- OD_2016 %>% mutate(ID = as.character(ID),
                           SchoolID = as.character(SchoolID)) %>%
  left_join(ttm_walk_2016 |> filter(travel_time_p50 <= 27), by=c("ID" = "from_id",
                                "SchoolID" = "to_id")) %>% 
  drop_na(travel_time_p50)

OD_2016 <- rbind(test,OD_2016) 
rm(test, ttm_car_2016, ttm_walk_2016)
```
```{r}
#reduce dimensions
OD_2016 <- OD_2016 %>% dplyr::select(-c("LUC1_DESC", "LUC2_DESC", "res_units", "DA_dwellings", "CMA_UID", "Region.Name", "geometry","PROP_AREA", "Area..sq.km."))

# Check the mean and median travel times by mode, the first is at the level of zn1259 and the second is at zn208
OD_2016 %>% 
  group_by(mode) %>% 
  summarize(mean_tt = mean(travel_time_p50),
            median_tt = median(travel_time_p50),
            min_tt = min(travel_time_p50),
            max_tt = max(travel_time_p50))
```

```{r}
#now let's add the communtiy names to the TAZ and summarize the % of students who use what mode by community. 
ggh_taz_communities <- st_intersection(st_centroid(ggh_taz) |> st_transform((st_crs(ggh_taz))), ham_community_bounds |> st_transform(st_crs(ggh_taz))) |> dplyr::select(c("GTA06", "REGION", "COMMUNITY_"))

od_trips_perc_2016 <- od_trips_perc_2016 %>% left_join(ggh_taz_communities |> st_drop_geometry() |> mutate(GTA06 = as.character(GTA06)), by=c("Origin"="GTA06"))

od_trips_perc_comm_2016 <- od_trips_perc_2016 %>% group_by(COMMUNITY_) %>%
  summarize(TAZOrig_car_trips = sum(TAZOrig_car_trips),
            TAZOrig_walk_trips = sum(TAZOrig_walk_trips)) %>% 
  mutate(TAZOrig_car_perc = TAZOrig_car_trips/(TAZOrig_car_trips + TAZOrig_walk_trips),
         COMMUNITY = case_when(COMMUNITY_ == "Ancaster" ~ "AN",
                                                         COMMUNITY_ == "Dundas" ~ "DU",
                                                         COMMUNITY_ == "Flamborough" ~ "FL",
                                                         COMMUNITY_ == "Glanbrook" ~ "GL",
                                                         COMMUNITY_ == "Hamilton" ~ "HA",
                                                         COMMUNITY_ == "Stoney Creek" ~ "SC")) |>
  dplyr::select(-c("COMMUNITY_"))

od_trips_perc_comm_2016
```

```{r}
# Lets add the % of car-mode used to the full OD list and then start transforming the frame for Spatial Availability calc.
OD_2016 <- OD_2016 |> left_join(od_trips_perc_comm_2016 |> dplyr::select(c("TAZOrig_car_perc", "COMMUNITY")), 
                     by = "COMMUNITY") %>%
  mutate(
  DA_pop_5to14 = ifelse(is.na(DA_pop_5to14), 0, DA_pop_5to14),
  DA_households = ifelse(is.na(DA_households), 0, DA_households))

#'mode_count' identifies if a OD pair (ID to SchoolID) is either just car (1) or car and walk (2). Join this to the full OD list.
OD_2016_mode_count <- OD_2016 |>
  group_by(ID, SchoolID) %>% 
  summarise(mode_count = n())

OD_2016 <- OD_2016 %>% left_join(OD_2016_mode_count, by = c("ID", "SchoolID"))
rm(OD_2016_mode_count)
```

```{r}
#now, here we make the assumption on how many students live in each parcel. So since we assume that any parcel could potentially go to any school - I think it makes sense to also assume that the most granular level of population data we have (DA-level) is equally assigned to each parcel and then equally assigned to each OD flow based on it's mode-split (on average ~80/20 car/walk).

#Count the number of ID's for each OD pair -- this can be seen as the number of 'trips' being made from a parcel. There are 147 elementary schools in 2011 -- all parcels can reach all schools by car in 60 mins. So if a parcel only has 1 mode_count (car) it will have a "DA_ID_count" of 147 (the number of OD home-to-all-school trips). This is the number that the rate of Students in DA per OD flows in the DA ("IDs_in_a_DA_count") is multipled by.
DA_ID_count <- OD_2016 |> group_by(ID) |> summarise(DA_ID_count = n()) 
OD_2016 <- OD_2016 |> left_join(DA_ID_count, by="ID")

#Here we calculate the number of OD trips per O (ID) that occur in each DA. This is how much the # of students in a DA is divded by for each OD. 
IDs_in_a_DA_count <- OD_2016 |> group_by(GeoUID) |> summarise(IDs_in_a_DA_count = sum(DA_ID_count)) 
OD_2016 <- OD_2016 |> left_join(IDs_in_a_DA_count, by="GeoUID")

#We calculate the _kid-pop_ per parcel and then we adjusted it based on if the OD flow is 'walking' or 'car'
OD_2016 <- OD_2016 |> mutate(kid_pop = DA_pop_5to14/IDs_in_a_DA_count*DA_ID_count, 
               kid_pop = ifelse(is.na(kid_pop) | is.nan(kid_pop) | kid_pop == Inf, 0, kid_pop),
               kid_popadj = case_when(mode == "w" & mode_count == 2 ~ (1.0000000-TAZOrig_car_perc)*2*kid_pop,
                                         mode == "c" & mode_count == 2 ~ (TAZOrig_car_perc)*2*kid_pop,
                                         mode == "c" & mode_count == 1 ~ kid_pop),
                  kid_popadj = ifelse(is.na(kid_popadj) | is.nan(kid_popadj) | kid_popadj == Inf, 0, kid_popadj))
```

```{r}
#checks!, these should be the total number of kids (^the code above) -- 58265.
OD_2016$kid_pop %>% sum()
OD_2016$kid_popadj %>% sum()
OD_2016 |> group_by(GeoUID) |> summarize(DA_pop_5to14 = mean(DA_pop_5to14)) |> dplyr::select(DA_pop_5to14) |> sum()
```

```{r}
#now add the impedance function, etc to the full OD 
OD_2016 <- OD_2016 |> mutate(
  travel_time_p50 = ifelse(travel_time_p50 == 0, 0.5, travel_time_p50),
  r = 1,
  f = ifelse(mode == "c", stats::dgamma(
    travel_time_p50, gamma2016_motor$estimate["shape"], gamma2016_motor$estimate["rate"]), 
    stats::dexp(
      travel_time_p50, exp2016_nonmotor$estimate["rate"])))
```

```{r}
#calculate spatial availability. 
avail_2016 <- sp_avail_detailed(OD_2016, ID, SchoolID, kid_popadj, OTGC2016, mode, r, f, alpha = 1)
save(avail_2016, file="data-products/avail_2016.RData")
save(OD_2016, file="data-products/OD_2016.RData")
rm(OD_2016)
```