### 2024-04-26 paper/Manuscript.qmd
Edits by Antonio

---
title: "School closures and consolidations: the active travel and emission implications of reduced accessibility"
bibliography: 
  - "`r system('kpsewhich ../bibliography/bibliography.bib', intern=TRUE)`"
linenumbers: true
numbersections: true
format:
  nature-pdf: default
editor: source
execute: 
  freeze: auto
author:
  - name: "Anastasia Soukhov"
    email: "soukhoa@mcmaster.ca"
    affiliation: McMaster University SEES
    footnote: "Corresponding Author"
  - name: "Christopher D. Higgins"
    email: "cd.higgins@utoronto.ca"
    affiliation: University of Toronto
  - name: "Moataz Mohamed"
    email: "mmohame@mcmaster.ca "
    affiliation: McMaster University CE
  - name: "Antonio Paez"
    email: "paezha@mcmaster.ca"
    affiliation: McMaster University SEES
address:
  - code: McMaster University SEES
    address: "School of Earth, Environment and Society, McMaster University, 1241 Main St. West, Hamilton, ON, L8S 4K1, Canada"
  - code: McMaster University CE
    address: "Department of Civil Engineering, McMaster University, 1241 Main St. West, Hamilton, ON, L8S 4K1, Canada"
  - code: University of Toronto
    address: "Department of Human Geography, University of Toronto Scarborough, 1265 Military Trail, Toronto, ON, M1C 1A4, Canada"  
abstract: | 
  | Reducing emissions and increasing active travel are often top priorities for many communities. These priorities, though, often are confronted with a pressure to reduce the fiscal burden of public services, including schools. Public schools have for some time now been pressured to do more with less, and governments are opting to close under-capacity schools and expand existing schools to accommodate lost capacity. The objective of this study is to quantifies the travel burdens of school consolidation/closure policy. We do this through the case of Hamilton, a mid-sized city in Ontario, Canada. Between the years 2011 and 2015, 7% of the elementary schools closed in the city. For this research, we quantify how these closures changed the accessibility landscape for the elementary school aged population, and the active travel and emission implications. In terms of methods, we use spatial availability, a singly-constrained multi-modal measure of spatial accessibility. Thanks to its proportional allocation mechanism, spatial availability allows for an interpretable comparison between 'before' and 'after' the implementation of school closure/consolidation policy. Analysis is conducted at the parcel-level with all school seats being eligible and travel behaviour based on observed home-to-school patterns for motorized and non-motorized modes. The system-wide impacts that result from the closures, such as potential GHG emission increases and walk minutes decreases, are estimated Overall, spatial availability decreased in the city. Though the school consolidation policy may have resulted in cost-savings in operation and maintenance of facilities, families, students, and the environment now pay the price.
keywords: [accessibility; spatial availability; carbon emissions; journey-to-school; walkability; active transportation; school consolidation policy; service provision thresholds]
---

```{r knitr-setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  cache = FALSE,
  warning = FALSE,
  message = FALSE,
  comment = '', 
  out.width = "1\\linewidth")
```

```{r, include=FALSE}
rm(list=ls())
```

```{r, include=FALSE}
library(dplyr) # A Grammar of Data Manipulation
library(ggplot2) # Create Elegant Data Visualisations Using the Grammar of Graphics
library(ggspatial) # Spatial Data Framework for ggplot2
library(here) # A Simpler Way to Find Your Files
library(patchwork) # The Composer of Plots
library(scales) # Scale Functions for Visualization
library(sf) # Simple Features for R
library(stplanr) # Sustainable Transport Planning
library(tidyverse) # Easily Install and Load the 'Tidyverse'
library(tmap) # Thematic Maps
library(units) # Measurement Units for R Vectors
```

```{r, include=FALSE}
load(file = paste0(here::here(),"/data-prep/data-products/cen_data_2011_hamiltononly.Rdata"))
cen_data_2011_hamiltononly <- cen_data_2011_1 |> filter(Region.Name == "Hamilton") #NOTE: GEOUID 35250242, 35250133, 35250457 do not have any child population
rm("cen_data_2011_1")
load(file=paste0(here::here(),"/data-prep/data-products/cen_data_2016_hamiltononly.Rdata"))
cen_data_2016_hamiltononly <- cen_data_2016_1 |> filter(Region.Name == "Hamilton") #NOTE: GEOUID 35250133, 35250398, 35250427 do not have any child population
rm("cen_data_2016_1")

ham_city_bound <- st_read(paste0(here::here(),"/data-prep/data-inputs/Boundaries/City_Boundary.shp"),
                          quiet = TRUE) |> st_transform(st_crs(cen_data_2011_hamiltononly))
ham_community_bounds <- st_read(paste0(here::here(),"/data-prep/data-inputs/Boundaries/Community_Boundaries.shp"),
                                quiet = TRUE) |> st_transform(st_crs(cen_data_2011_hamiltononly))
```

```{r, include=FALSE}
cen_data_2011_hamiltononly <- 
  cen_data_2011_hamiltononly %>% dplyr::select( #removing the variables we don't need for this analysis
    -c("Shape.Area", "Quality.Flags", "Type", "CD_UID", "NHS.Non.Return.Rate", "CSD_UID", "CT_UID", "CT_UID","NHS.Non.Return.Rate.1", "v_CA11N_2564..Median.after.tax.household.income..", "v_CA11F_17..15.to.19.years")) %>%
  rename( #renaming the variables
    "DA_households" = "Households",
    "DA_dwellings" = "Dwellings",
    "DA_tot_pop" = "Population",
    "DA_med_AT_family_income" = "v_CA11N_2458..Median.after.tax.family.income..",
    "DA_med_AT_lone-family_income" = "v_CA11N_2476..Median.after.tax.family.income..",
    "DA_prev_LIMAT_under18" = "v_CA11N_2609..Less.than.18.years..",
    "DA_pop_5to9" = "v_CA11F_11..5.to.9.years",
    "DA_pop_10to14" = "v_CA11F_14..10.to.14.years") %>%
  mutate(DA_pop_5to14 = DA_pop_5to9+(DA_pop_10to14)) |> #*0.9)) %>%
  dplyr::select(-c("DA_pop_5to9","DA_pop_10to14"))

cen_data_2016_hamiltononly <- 
  cen_data_2016_hamiltononly %>% dplyr::select( #removing the variables we don't need for this analysis
    -c("Shape.Area",  "Type", "CD_UID", "CSD_UID", "CT_UID", "CT_UID", "v_CA16_2398..Median.after.tax.income.of.households.in.2015....", "v_CA16_64..15.to.19.years")) %>%
  rename( #renaming the variables
    "DA_households" = "Households",
    "DA_dwellings" = "Dwellings",
    "DA_tot_pop" = "Population",
    "DA_med_AT_family_income" = "v_CA16_2448..Median.after.tax.income.of.economic.families.in.2015....",
    "DA_med_AT_lone-family_income" = "v_CA16_2460..Median.after.tax.income.of.lone.parent.economic.families.in.2015....",
    "DA_prev_LIMAT_under18" = "v_CA16_2543..0.to.17.years....",
    "DA_pop_5to9" = "v_CA16_25..5.to.9.years",
    "DA_pop_10to14" = "v_CA16_43..10.to.14.years") %>% mutate(DA_pop_5to14 = DA_pop_5to9+(DA_pop_10to14)) |> #*0.9)) %>%
    dplyr::select(-c("DA_pop_5to9","DA_pop_10to14"))
```

```{r, include=FALSE}
library(hamONdests) # A R package of Hamilton, Ontario destinations
data(Schools_2011_2016)
ALL_SCHOOLS_Elem <- Schools_2011_2016 %>% filter(Level == "Elementary")
rm("Schools_2011_2016")
data(School_Catchments_2011_2016)
ALL_SCHOOLS_CATCH_Elem <- School_Catchments_2011_2016 %>% filter(Level == "Elementary")
rm("School_Catchments_2011_2016")

load(file = paste0(here::here(),"/data-prep/data-products/TTS_gamma2011_motor.Rdata"))
load(file = paste0(here::here(),"/data-prep/data-products/TTS_gamma2016_motor.Rdata"))
load(file = paste0(here::here(),"/data-prep/data-products/TTS_exp2011_nonmotor.Rdata"))
load(file = paste0(here::here(),"/data-prep/data-products/TTS_exp2016_nonmotor.Rdata"))
```

```{r}
# # maybe consider using year 2021 as a comparision? 2011 vs 2016 vs 2021?
# SCHOOLS_2023 <- st_read("data-prep/data-inputs/Educational_Institutions-2023/Educational_Institutions.shp") |> filter(CATEGORY == "Elementary School" & (SCHOOL_BOA == "Hamilton-Wentworth Catholic District School Board" | SCHOOL_BOA == "Hamilton-Wentworth District School Board" )) |> st_transform(st_crs(cen_data_2011_hamiltononly))
```

```{r augmenting-catchment, include=FALSE}
ALL_SCHOOLS_CATCH_Elem_extra2011 <- ALL_SCHOOLS_CATCH_Elem |>
  left_join(ALL_SCHOOLS_Elem |> st_drop_geometry() |> group_by(CID2011) |> 
              summarize(OTGC = sum(OTGC2011, na.rm = T)), 
            by = c("CID" = "CID2011")) |> filter(!is.na(OTGC))

ALL_SCHOOLS_CATCH_Elem_extra2016 <- ALL_SCHOOLS_CATCH_Elem |>
  left_join(ALL_SCHOOLS_Elem |> st_drop_geometry() |> group_by(CID2016) |> 
              summarize(OTGC = sum(OTGC2016, na.rm = T)), 
            by = c("CID" = "CID2016")) |> filter(!is.na(OTGC))

ALL_SCHOOLS_CATCH_Elem_extra <- rbind(ALL_SCHOOLS_CATCH_Elem_extra2011, ALL_SCHOOLS_CATCH_Elem_extra2016)
rm(ALL_SCHOOLS_CATCH_Elem_extra2011,ALL_SCHOOLS_CATCH_Elem_extra2016)

#checks
sum(ALL_SCHOOLS_Elem$OTGC2011, na.rm=T) + sum(ALL_SCHOOLS_Elem$OTGC2016, na.rm=T)
ALL_SCHOOLS_CATCH_Elem_extra$OTGC |> sum(na.rm=T)

#reformating schools for mapping
ALL_SCHOOLS_Elem_nochangeorexpanded <- ALL_SCHOOLS_Elem |> filter(Year == "2011 and 2016") |> mutate(Year = ifelse(Year == "2011 and 2016", "2011", Year))
ALL_SCHOOLS_Elem_formapping <- rbind(ALL_SCHOOLS_Elem, ALL_SCHOOLS_Elem_nochangeorexpanded)
rm(ALL_SCHOOLS_Elem_nochangeorexpanded)
ALL_SCHOOLS_Elem_formapping <- ALL_SCHOOLS_Elem_formapping |> mutate(Year = ifelse(Year == "2011 and 2016", "2016", Year))
```

```{r descriptive-catchment-schools-plot, eval=FALSE}
tmap_defaults <-
  tm_scale_bar(position = c("LEFT", "BOTTOM"), breaks = c(0, 5, 10, 15))+
  tm_compass(size = 1.5, position = c("right", 'top')) +
  tm_layout(legend.position = c("left", "top")) #legend.hist.size = 10,

catchment_schools_plot <-
  tm_shape(ALL_SCHOOLS_CATCH_Elem_extra) +
  tm_polygons(col = "OTGC", border.alpha=0.1, palette = "Greys",
              labels = c("[181-318)", "[318-406)", "[406-566)","[566-1417]"),
              breaks = c(181, 318, 406, 566, 1417),
              title = "On the ground\ncapacity (OTGC)")+
  tm_facets(by=c("System","Year"))+
  tm_shape(ALL_SCHOOLS_Elem_formapping) +
  tm_dots(col = "Status", size=0.1, title = "School status between\n2011 and 2016",
          palette = c("#ffd14d", "black", "darkgreen","#0E86D4","#90EE90"),
          labels = c("No change","Closed before 2016", "Expanded before 2016","Moved before 2016","New after 2011")          
  ) +
  tm_facets(by=c("Year","System"))+
  tmap_defaults +
  tm_layout(panel.labels = list(c('HWDCSB (Catholic Public)','HWDSB (Public)'),c('2011','2016')))

tmap_save(catchment_schools_plot, paste0(here::here(),"/Figures/catchment_schools_plot.png"), height=7, width=12, dpi=600)
```

```{r,include=FALSE}
load(file = paste0(here::here(),"/data-prep/data-products/avail_2011.RData"))
load(file = paste0(here::here(),"/data-prep/data-products/avail_2016.RData"))
```

```{r}
community_car_perc_2011 <- avail_2011 |>
  group_by(COMMUNITY) |> 
  summarize(TAZOrig_car_perc = mean(TAZOrig_car_perc, na.rm=T))

community_car_perc_2016 <- avail_2016 |> group_by(COMMUNITY) |> summarize(TAZOrig_car_perc = mean(TAZOrig_car_perc, na.rm=T))
```

```{r, eval=FALSE}
#check to make sure that V_i is equal to the total capacity in Hamilton. It is!
avail_2011$V_ij %>% sum()
ALL_SCHOOLS_Elem$OTGC2011 %>% sum(na.rm = TRUE)

avail_2016$V_ij %>% sum()
ALL_SCHOOLS_Elem$OTGC2016 %>% sum(na.rm = TRUE)

#this should equal 58,265, still does -- great!
avail_2011 %>%
  group_by(ID) %>%
  summarize(kid_popadj = sum(kid_popadj)) |> dplyr::select(kid_popadj) %>% sum()

#this should equal 58,865, still does -- great!
avail_2016 %>%
  group_by(ID) %>%
  summarize(kid_popadj = sum(kid_popadj)) |> dplyr::select(kid_popadj) %>% sum()
```

```{r}
SAvail_im_DA2011 <- avail_2011 %>%
  group_by(GeoUID, 
           mode) %>%
  summarize(COMMUNITY = first(COMMUNITY),
            total_tt = sum(travel_time_p50*kid_popadj),
            V_im = sum(V_ij),
            kid_popadj_m = sum(kid_popadj),
            DA_pop_5to14 = mean(DA_pop_5to14, na.rm=T),
            DA_prev_LIMAT_under18 = mean(DA_prev_LIMAT_under18),
            .groups = "drop") |>
  mutate(emissions_total_car_tt = ifelse(mode == "c", 0.1*total_tt, 0),
         v_im_kid_popadj = V_im/kid_popadj_m,
         v_im_kid_popadj_LIMAT = v_im_kid_popadj*(DA_prev_LIMAT_under18/100),
         v_im_kid_popadj_LIMAT = ifelse(is.na(v_im_kid_popadj_LIMAT) | is.nan(v_im_kid_popadj_LIMAT), 0, v_im_kid_popadj_LIMAT),
         V_i_kid_popadj_LIMAT = V_im*(DA_prev_LIMAT_under18/100),
         V_i_kid_popadj_LIMAT = ifelse(is.na(V_i_kid_popadj_LIMAT) | is.nan(V_i_kid_popadj_LIMAT), 0, V_i_kid_popadj_LIMAT)) |>
  filter(GeoUID != 35250242 & GeoUID != 35250133 & GeoUID != 35250457 ) #these GeoUID's have 0 kid population, so let's drop so we don't have zeros

SAvail_im_DA2016 <- avail_2016 %>%
  group_by(GeoUID, 
           mode) %>%
  summarize(COMMUNITY = first(COMMUNITY),
            total_tt = sum(travel_time_p50*kid_popadj),
            V_im = sum(V_ij),
            kid_popadj_m = sum(kid_popadj),
            DA_pop_5to14 = mean(DA_pop_5to14, na.rm=T),
            DA_prev_LIMAT_under18 = mean(DA_prev_LIMAT_under18),
            .groups = "drop") |>
  mutate(emissions_total_car_tt = ifelse(mode == "c", 0.1*total_tt, 0),
         v_im_kid_popadj = V_im/kid_popadj_m,
         v_im_kid_popadj = ifelse(is.na(v_im_kid_popadj) | is.nan(v_im_kid_popadj), 0, v_im_kid_popadj),
         v_im_kid_popadj_LIMAT = v_im_kid_popadj*(DA_prev_LIMAT_under18/100),
         v_im_kid_popadj_LIMAT = ifelse(is.na(v_im_kid_popadj_LIMAT) | is.nan(v_im_kid_popadj_LIMAT), 0, v_im_kid_popadj_LIMAT),
         V_i_kid_popadj_LIMAT = V_im*(DA_prev_LIMAT_under18/100),
         V_i_kid_popadj_LIMAT = ifelse(is.na(V_i_kid_popadj_LIMAT) | is.nan(V_i_kid_popadj_LIMAT), 0, V_i_kid_popadj_LIMAT)) |>
  filter(GeoUID != 35250398 & GeoUID != 35250133 & GeoUID != 35250427 ) #these GeoUID's have 0 kid population, so let's drop so we don't have zeros

#add the geometries -- census DAs
SAvail_im_DA2011 <- SAvail_im_DA2011 |>
  left_join(cen_data_2011_hamiltononly |> filter(Region.Name == "Hamilton") |> dplyr::select(GeoUID), by="GeoUID") |> st_set_geometry("geometry")

SAvail_im_DA2016 <- SAvail_im_DA2016 |>
  left_join(cen_data_2016_hamiltononly |> filter(Region.Name == "Hamilton") |> dplyr::select(GeoUID), by="GeoUID") |> st_set_geometry("geometry")
```
```{r}
#Create 1 sf with data joined, no changes
SAvail_im_DA <-
  rbind(SAvail_im_DA2011 |> mutate(year ="2011"),
        SAvail_im_DA2016 |> mutate(year ="2016")) |>
  mutate(mode = case_when(mode == "c" ~ "motorized",
                          mode == "w" ~ "non-motorized"))

#calc changes between the years - make that 1 sf
SAvail_im_DA_changes <- full_join(SAvail_im_DA2011|>st_drop_geometry(),
                                  SAvail_im_DA2016, by=c("GeoUID", "mode")) |>
  transmute(GeoUID,
            mode = case_when(mode == "c" ~ "motorized",
                             mode == "w" ~ "non-motorized"),
            COMMUNITY = dplyr::coalesce(COMMUNITY.x, COMMUNITY.y),
            c_total_tt = total_tt.y - total_tt.x, #2016 minus 2011 tt
            c_V_im = V_im.y -V_im.x,
            c_V_im_perc = (V_im.y -V_im.x)/V_im.x,
            c_kid_popadj_m = kid_popadj_m.y - kid_popadj_m.x,
            c_DA_pop_5to14 = DA_pop_5to14.y - DA_pop_5to14.x,
            c_DA_prev_LIMAT_under18 = DA_prev_LIMAT_under18.y - DA_prev_LIMAT_under18.x,
            c_emissions_total_car_tt = emissions_total_car_tt.y - emissions_total_car_tt.x,
            c_v_im_kid_popadj = v_im_kid_popadj.y - v_im_kid_popadj.x,
            c_v_im_kid_popadj_perc = (v_im_kid_popadj.y - v_im_kid_popadj.x)/v_im_kid_popadj.x,
            c_v_im_kid_popadj_LIMAT = v_im_kid_popadj_LIMAT.y - v_im_kid_popadj_LIMAT.x,
            c_V_i_kid_popadj_LIMAT = V_i_kid_popadj_LIMAT.y - V_i_kid_popadj_LIMAT.x,
            geometry) |> st_as_sf()

```

```{r}
#repeat but for all 1 mode
SAvail_i_DA2011 <- avail_2011 %>%
  group_by(GeoUID) %>%
  summarize(COMMUNITY = first(COMMUNITY),
            total_tt = sum(travel_time_p50*kid_popadj, na.rm = T),
            V_i = sum(V_ij, na.rm = T),
            DA_pop_5to14 = mean(DA_pop_5to14, na.rm=T),
            kid_pop_DAavg = mean(kid_pop, na.rm = T),
            DA_prev_LIMAT_under18 = mean(DA_prev_LIMAT_under18, na.rm = T),
            .groups = "drop") |>
  filter(GeoUID != 35250242 & GeoUID != 35250133 & GeoUID != 35250457 ) #these GeoUID's have 0 kid population, so let's drop so we don't have zeros

SAvail_i_DA2016 <- avail_2016 %>%
  group_by(GeoUID) %>%
  summarize(COMMUNITY = first(COMMUNITY),
            total_tt = sum(travel_time_p50*kid_popadj),
            V_i = sum(V_ij, na.rm = T),
            DA_pop_5to14 = mean(DA_pop_5to14, na.rm=T),
            kid_pop_DAavg = mean(kid_pop, na.rm = T),
            DA_prev_LIMAT_under18 = mean(DA_prev_LIMAT_under18, na.rm = T),
            .groups = "drop") |>
  filter(GeoUID != 35250398 & GeoUID != 35250133 & GeoUID != 35250427 ) #these GeoUID's have 0 kid population, so let's drop so we don't have zeros

#add the geometries -- census DAs
SAvail_i_DA2011 <- SAvail_i_DA2011 |>
  left_join(cen_data_2011_hamiltononly |> filter(Region.Name == "Hamilton") |> dplyr::select(GeoUID), by="GeoUID") |> mutate(Year = "2011") |> st_set_geometry("geometry")

SAvail_i_DA2016 <- SAvail_i_DA2016 |>
  left_join(cen_data_2016_hamiltononly |> filter(Region.Name == "Hamilton") |> dplyr::select(GeoUID), by="GeoUID") |> mutate(Year = "2016") |> st_set_geometry("geometry")

#join them together!
SAvail_i_DA_extra <- rbind(SAvail_i_DA2011, SAvail_i_DA2016)
rm(SAvail_i_DA2011,SAvail_i_DA2016)
```

```{r}
#now do this at the community level
ham_community_bounds <- ham_community_bounds |> 
  mutate(COMMUNITY = case_when(COMMUNITY_ == "Ancaster" ~ "AN",
                               COMMUNITY_ == "Dundas" ~ "DU",
                               COMMUNITY_ == "Flamborough" ~ "FL",
                               COMMUNITY_ == "Glanbrook" ~ "GL",
                               COMMUNITY_ == "Hamilton" ~ "HA",
                               COMMUNITY_ == "Stoney Creek" ~ "SC"))

SAvail_im_COMM2011 <- SAvail_im_DA2011 %>% st_drop_geometry() |>
  group_by(COMMUNITY, 
           mode) %>%
  summarize(total_tt = sum(total_tt, na.rm=T),
            V_im = sum(V_im, na.rm=T),
            kid_popadj_m = sum(kid_popadj_m, na.rm=T),
            COMM_prev_LIMAT_under18 = mean(DA_prev_LIMAT_under18, na.rm=T),
            emissions_total_car_tt = sum(emissions_total_car_tt, na.rm=T),
            .groups = "drop") |>
  mutate(v_im_kid_popadj = V_im/kid_popadj_m,
         v_im_kid_popadj_LIMAT = v_im_kid_popadj*(COMM_prev_LIMAT_under18/100),
         v_im_kid_popadj_LIMAT = ifelse(is.na(v_im_kid_popadj_LIMAT) | is.nan(v_im_kid_popadj_LIMAT), 0, v_im_kid_popadj_LIMAT),
         V_i_kid_popadj_LIMAT = V_im*(COMM_prev_LIMAT_under18/100),
         V_i_kid_popadj_LIMAT = ifelse(is.na(V_i_kid_popadj_LIMAT) | is.nan(V_i_kid_popadj_LIMAT), 0, V_i_kid_popadj_LIMAT))  |> left_join(ham_community_bounds |> dplyr::select(-c("OBJECTID")), by=c("COMMUNITY")) |> st_set_geometry("geometry")

SAvail_im_COMM2016 <- SAvail_im_DA2016 %>% st_drop_geometry() |>
  group_by(COMMUNITY, 
           mode) %>%
  summarize(total_tt = sum(total_tt, na.rm=T),
            V_im = sum(V_im, na.rm=T),
            kid_popadj_m = sum(kid_popadj_m, na.rm=T),
            COMM_prev_LIMAT_under18 = mean(DA_prev_LIMAT_under18, na.rm=T),
            emissions_total_car_tt = sum(emissions_total_car_tt, na.rm=T),
            .groups = "drop") |>
  mutate(v_im_kid_popadj = V_im/kid_popadj_m,
         v_im_kid_popadj_LIMAT = v_im_kid_popadj*(COMM_prev_LIMAT_under18/100),
         v_im_kid_popadj_LIMAT = ifelse(is.na(v_im_kid_popadj_LIMAT) | is.nan(v_im_kid_popadj_LIMAT), 0, v_im_kid_popadj_LIMAT),
         V_im_kid_popadj_LIMAT = V_im*(COMM_prev_LIMAT_under18/100),
         V_im_kid_popadj_LIMAT = ifelse(is.na(V_im_kid_popadj_LIMAT) | is.nan(V_im_kid_popadj_LIMAT), 0, V_im_kid_popadj_LIMAT))  |> left_join(ham_community_bounds |> dplyr::select(-c("OBJECTID")), by=c("COMMUNITY")) |> st_set_geometry("geometry")
```

```{r}
#repeat above but with just i, this is to get the correct LIM value.
SAvail_i_COMM2011 <- SAvail_im_DA2011 %>% st_drop_geometry() |>
  group_by(COMMUNITY) %>%
  summarize(COMM_prev_LIMAT_under18 = mean(DA_prev_LIMAT_under18, na.rm=T),
            .groups = "drop")

SAvail_i_COMM2016 <- SAvail_im_DA2016 %>% st_drop_geometry() |>
  group_by(COMMUNITY) %>%
  summarize(COMM_prev_LIMAT_under18 = mean(DA_prev_LIMAT_under18, na.rm=T),
            .groups = "drop")

#calc changes between the years - make that 1 sf
SAvail_i_COMM_changes <- full_join(SAvail_i_COMM2011 |> st_drop_geometry(),
                                   SAvail_i_COMM2016, by=c("COMMUNITY")) |>
  transmute(COMMUNITY,
            prev_LIMAT_2011 = COMM_prev_LIMAT_under18.x,
            prev_LIMAT_2016 = COMM_prev_LIMAT_under18.y,
            c_prev_LIMAT_perc = ((COMM_prev_LIMAT_under18.y - COMM_prev_LIMAT_under18.x)/
                                   COMM_prev_LIMAT_under18.x )|> scales::percent())
```

```{r}
#calc changes between the years - make that 1 sf
SAvail_im_COMM_changes <- full_join(SAvail_im_COMM2011|>st_drop_geometry(),
                                    SAvail_im_COMM2016, by=c("COMMUNITY", "mode")) |>
  transmute(COMMUNITY,
            mode = case_when(mode == "c" ~ "motorized",
                             mode == "w" ~ "non-motorized"),
            COMMUNITY_ = dplyr::coalesce(COMMUNITY_.x, COMMUNITY_.y),
            kid_2011 = kid_popadj_m.x,
            kid_2016 = kid_popadj_m.y,
            c_kid_perc = ((kid_popadj_m.y - kid_popadj_m.x)/kid_popadj_m.x) |> scales::percent(),
            kid_proportion_2011 = (kid_2011/sum(kid_2011)),
            kid_proportion_2016 = (kid_2016/sum(kid_2016)),
            c_kid_proportion_perc = ((kid_proportion_2016 - kid_proportion_2011)/kid_proportion_2011)
            |> scales::percent(),
            kid_proportion_2011_perc = (kid_proportion_2011) |> scales::percent(),
            kid_proportion_2016_perc = (kid_proportion_2016) |> scales::percent(),
            V_im_2011 = V_im.x,
            V_im_2016 = V_im.y,
            c_V_im_perc = ((V_im.y -V_im.x)/V_im.x) |> scales::percent(),
            V_im_proportion_2011 = (V_im.x/sum(V_im.x)),
            V_im_proportion_2016 = (V_im.y/sum(V_im.y)),
            c_V_im_proportion_perc = ((V_im_proportion_2016 - V_im_proportion_2011)/V_im_proportion_2011)
            |> scales::percent(),
            V_im_proportion_2011_perc = (V_im_proportion_2011) |> scales::percent(),
            V_im_proportion_2016_perc = (V_im_proportion_2016) |> scales::percent(),
            v_im_2011 = v_im_kid_popadj.x,
            v_im_2016 = v_im_kid_popadj.y,
            c_v_im_perc = ((v_im_kid_popadj.y - v_im_kid_popadj.x)/v_im_kid_popadj.x) |> scales::percent(),
            total_tt_2011 = total_tt.x,
            total_tt_2016 = total_tt.y,
            total_tt_proportion_2011 = (total_tt.x/sum(total_tt.x)),
            total_tt_proportion_2016 = (total_tt.y/sum(total_tt.y)),
            c_total_tt_proportion_perc = ((total_tt_proportion_2016 - total_tt_proportion_2011)/total_tt_proportion_2011)
            |> scales::percent(),
            total_tt_proportion_2011 = (total_tt_proportion_2011) |> scales::percent(),
            total_tt_proportion_2016 = (total_tt_proportion_2016) |> scales::percent(),
            c_total_tt_perc = ((total_tt.y - total_tt.x)/
                                 total_tt.x)|> scales::percent(),
            total_tt_perkid_2011 = total_tt.x/kid_2011,
            total_tt_perkid_2016 = total_tt.y/kid_2016,
            c_total_ttperkid_perc = ((total_tt_perkid_2016 - total_tt_perkid_2011)/
                                       total_tt_perkid_2011)|> scales::percent(),
            emissions_2011 = emissions_total_car_tt.x,
            emissions_2016 = emissions_total_car_tt.y,
            c_emissions_perc = ((emissions_total_car_tt.y - emissions_total_car_tt.x)/
                                  emissions_total_car_tt.x)|> scales::percent(),
            emissions_proportion_2011 = (emissions_2011/sum(emissions_2011)),
            emissions_proportion_2016 = (emissions_2016/sum(emissions_2016)),
            c_emissions_proportion = ((emissions_proportion_2016 - emissions_proportion_2011)/emissions_proportion_2011)
            |> scales::percent(),
            emissions_proportion_2011_perc = (emissions_proportion_2011) |> scales::percent(),
            emissions_proportion_2016_perc = (emissions_proportion_2016) |> scales::percent(),
            emissions_perkid_2011 = emissions_total_car_tt.x/kid_popadj_m.x,
            emissions_perkid_2016 = emissions_total_car_tt.y/kid_popadj_m.y,
            c_emissions_perkid_perc = ((emissions_perkid_2016 - emissions_perkid_2011)/
                                         emissions_perkid_2011)|> scales::percent(),
            geometry) |> 
  left_join(SAvail_i_COMM_changes, by="COMMUNITY") |> st_as_sf() 
```

```{r student-popabs-lowincperc-poprate-plot, eval=FALSE}
tmap_defaults <-
  tm_scale_bar(position = c("LEFT", "BOTTOM"), breaks = c(0, 5, 10, 15))+
  tm_compass(size = 1.5, position = c("right", 'top')) +
  tm_layout(legend.position = c("left", "top")) #legend.hist.size = 10,

student_pop_absolute_plot <-
  tm_shape(SAvail_i_DA_extra |> mutate(extra_col_for_panel = "HAMILTON")) +
  tm_polygons(col = "DA_pop_5to14", border.alpha=0, 
              palette = "Greys",
              labels = c("[0-40)", "[40-50)", "[50-75)","[75-835]"),
              breaks = c(0, 40, 50, 75, 835),
              title = "Population aged 5 to 14"
  )+
  tm_facets(by=c("Year"), ncol=2, nrow=1)+
  tmap_defaults +
  tm_layout(panel.label.height = 0.8)

student_pop_rate_plot <-
  tm_shape(SAvail_i_DA_extra) +
  tm_polygons(col = "kid_pop_DAavg", border.alpha=0, 
              palette = "Blues",
              labels = c("[0-0.0018)", "[0.0018-0.0025)", "[0.0025-0.0038)","[0.0038-0.59]"),
              breaks = c(0, 0.0018, 0.0025, 0.0038, 0.60),
              title = "Average pop. aged 5 to 14\nper residential unit"
  )+
  tm_facets(by=c("Year"), ncol=2, nrow=1)+
  tmap_defaults +
  tm_layout(panel.show=FALSE)

LIMAT_under18_DA_plot <-
  tm_shape(SAvail_i_DA_extra) +
  tm_polygons(col = "DA_prev_LIMAT_under18", border.alpha=0, 
              palette = "Reds",
              labels = c("[0-4.3)", "[4.3-13.0)", "[13.0-30.8)","[30.8-100)"),
              breaks = c(0, 4.30, 13.00, 30.8, 150),
              textNA = "Unavailable",
              title = "Proportion of LIM-AT households\n with children < 18"
  )+
  tm_facets(by=c("Year"), ncol=2, nrow=1)+
  tmap_defaults  +
  tm_layout(panel.show=FALSE)

stud_pop_rate_LIM_plot <- tmap_arrange(student_pop_absolute_plot, student_pop_rate_plot, LIMAT_under18_DA_plot, nrow = 3)

tmap_save(stud_pop_rate_LIM_plot, paste0(here::here(),"/Figures/stud_pop_rate_LIM_plot.png"), height=7, width=12, dpi=600)
```

```{r importing-taz-and-associated-flows, include=FALSE}
#Read the boundaries of the traffic analysis zones (retrieved from the TTS 2016):
ggh_taz <- st_read(paste0(here::here(), "/data-prep/data-inputs/TTS/tts06_83_region.shp"),
                   quiet = TRUE) %>% 
  st_transform(crs = 32617)

# Prepare traffic analysis zones:
ggh_taz <- ggh_taz %>%
  transmute(GTA06 = as.character(GTA06),
            AREA = st_area(geometry) %>% 
              set_units(km^2) %>% 
              drop_units())
#Extract the IDs for all TAZ:
ghta_taz_id <- ggh_taz$GTA06

#Read the boundaries of the traffic analysis zones (retrieved from the TTS 2016):
ggh_pd <- st_read(paste0(here::here(), "/data-prep/data-inputs/TTS/tts06_pd_83.shp"),
                  quiet = TRUE) %>%
  st_transform(crs = 32617)

#Adding region names to the object as they are not included in the origin data: (http://dmg.utoronto.ca/pdf/tts/2016/2016TTS_Conduct.pdf)

ggh_pd <- ggh_pd %>%
  mutate(REGION_name = 
           ifelse(REGION == 1, "Toronto", 
                  ifelse(REGION == 2, "Durham", 
                         ifelse(REGION == 3, "York", 
                                ifelse(REGION == 4, "Peel", 
                                       ifelse(REGION == 5, "Halton", 
                                              ifelse(REGION == 6, "Hamilton", 
                                                     ifelse(REGION == 11, "Niagara", 
                                                            ifelse(REGION == 12, "Waterloo", 
                                                                   ifelse(REGION == 13, "Guelph", 
                                                                          ifelse(REGION == 14, "Wellington", 
                                                                                 ifelse(REGION == 15, "Orangeville", 
                                                                                        ifelse(REGION == 16, "Barrie", 
                                                                                               ifelse(REGION == 17, "Simcoe", 
                                                                                                      ifelse(REGION == 18, "Kawartha Lakes", 
                                                                                                             ifelse(REGION == 19, "Peterborough City", 
                                                                                                                    ifelse(REGION == 20, "Peterborough County", 
                                                                                                                           ifelse(REGION == 21, "Orillia", 
                                                                                                                                  ifelse(REGION == 22, "Dufferin", 
                                                                                                                                         ifelse(REGION == 23, "Brantford", 
                                                                                                                                                ifelse(REGION == 24, "Brant", "NOOOO")))))))))))))))))))))

# add 'REGION' from the planning distract object to the `ggh_taz`. 
#buffering to clean up the edges of these objects -> there wre issues using st_join. 
ggh_taz <- st_buffer(ggh_taz, dist=0)
ggh_pd <- st_buffer(ggh_pd, dist=0)

ggh_taz <- st_join(ggh_taz, ggh_pd %>% transmute(REGION),left = TRUE, largest = TRUE)
ham_taz <- ggh_taz %>% dplyr::filter(REGION==6) #Filter out everything other than Planning District 6 = Hamilton
```

```{r include=FALSE}
#the tts origin-destination trips, by mode
load(file = paste0(here::here(),"/data-prep/data-inputs/TTS/od_trips_perc_2011.Rdata"))
load(file = paste0(here::here(),"/data-prep/data-inputs/TTS/od_trips_perc_2016.Rdata"))

ham_taz_centroids <- ham_taz |> st_centroid() |> dplyr::select(GTA06)

od_trips_perc_car <- rbind(
  rbind(od_trips_perc_2011 |> group_by(Origin) |> 
          summarize(TAZOrig_car_perc = 
                      sum(TAZOrig_car_trips, na.rm = T)/sum(TAZOrig_walk_trips+TAZOrig_car_trips, na.rm = T)) |>
          mutate(year = "2011")),
  od_trips_perc_2016 |> group_by(Origin) |> 
    summarize(TAZOrig_car_perc = 
                sum(TAZOrig_car_trips, na.rm = T)/sum(TAZOrig_walk_trips+TAZOrig_car_trips, na.rm = T)) |>
    mutate(year = "2016"))

od_trips_perc_car <- left_join(ham_taz,od_trips_perc_car, by=c("GTA06"="Origin"))

od_trips_perc_walk <- rbind(
  rbind(od_trips_perc_2011 |> group_by(Origin) |> 
          summarize(TAZOrig_walk_perc = 
                      sum(TAZOrig_walk_trips)/sum(TAZOrig_walk_trips+TAZOrig_car_trips)) |>
          mutate(year = "2011")),
  od_trips_perc_2016 |> group_by(Origin) |> 
    summarize(TAZOrig_walk_perc = 
                sum(TAZOrig_walk_trips)/sum(TAZOrig_walk_trips+TAZOrig_car_trips)) |>
    mutate(year = "2016"))

od_trips_perc_walk <- left_join(ham_taz, od_trips_perc_walk, by=c("GTA06"="Origin"))

desire_lines_ham_2011 <- od2line(od_trips_perc_2011, ham_taz_centroids)
desire_lines_ham_2011 <- desire_lines_ham_2011 |> mutate(all = TAZOrig_car_trips + TAZOrig_walk_trips,
                                                         Year = "2011")

desire_lines_ham_2016 <- od2line(od_trips_perc_2016, ham_taz_centroids)
desire_lines_ham_2016 <- desire_lines_ham_2016 |> mutate(all = TAZOrig_car_trips + TAZOrig_walk_trips,
                                                         Year = "2016")

desire_lines_ham <- rbind(desire_lines_ham_2011, desire_lines_ham_2016)
rm(desire_lines_ham_2011, desire_lines_ham_2016)
```

```{r ploting-od-taz-flows, eval=FALSE}
car_OD_flow_plot <-  tm_shape(od_trips_perc_car |> mutate(mode = "motorized"), bbox=SAvail_i_DA_extra) +
  tm_polygons(col = "TAZOrig_car_perc", border.alpha = 0.4, border.col = "grey20", lwd=0.2,
              palette = "Greys",
              breaks = c(0, .25, .50, .75, 1),
              title = "Proportion of mode use",
              textNA = "Unavailable",
              colorNA = "white")+
  #tm_shape(SAvail_i_DA_extra) + tm_borders(alpha = 0.4, col = "grey20", lwd=0.2)+
  tm_shape(ham_taz) + tm_borders(alpha=0.4, col = "grey40", lwd=0.2) +
  tm_shape(ham_community_bounds) + tm_borders(alpha = 0.8, col = "lightblue", lwd=2)+
  tm_shape(ham_taz_centroids) + tm_dots(alpha=0.5, col="darkblue", size=0.01) +
  tm_shape(desire_lines_ham |> mutate(mode = "motorized")) +
  tm_lines(palette = "Reds", breaks = c(6, 22, 26, 49, 353),
           title.col = "Number of trips (motorized)",
           col = "TAZOrig_car_trips",
           alpha = 0.9,
           scale = 3,
           lwd = "all",
           legend.lwd.show = FALSE) +
  tm_facets(by = c("mode","Year"), nrow=1)+
  tmap_defaults +
  tm_layout(panel.label.height = 0.8)

walk_OD_flow <- tm_shape(od_trips_perc_walk |> mutate(mode = "non-motorized"), bbox=SAvail_i_DA_extra) +
  tm_polygons(col = "TAZOrig_walk_perc", border.alpha = 0.4, border.col = "grey20", lwd=0.2,
              palette = "Greys",
              breaks = c(0, .25, .50, .75, 1),
              title = "Proportion of mode use",
              textNA = "Unavailable",
              colorNA = "white",
              legend.show=FALSE)+
  #tm_shape(SAvail_i_DA_extra) + tm_borders(alpha = 0.4, col = "cornflowerblue", lwd=0.2)+
  tm_shape(ham_taz) + tm_borders(alpha=0.4, col = "grey40", lwd=0.2) +
  tm_shape(ham_community_bounds) + tm_borders(alpha = 0.8, col = "lightblue", lwd=2)+
  tm_shape(ham_taz_centroids) + tm_dots(alpha=0.5, col="darkblue", size=0.01) +
  tm_shape(desire_lines_ham |> filter(TAZOrig_walk_trips >= 1) 
           |> mutate(mode = "non-motorized")) +
  tm_lines(palette = "Greens", breaks = c(12, 23, 44, 67, 174),
           title.col = "Number of trips (non-motorized)",
           col = "TAZOrig_walk_trips",
           alpha = 0.9,
           scale = 4,
           lwd = "all",
           legend.lwd.show = FALSE) +
  tm_facets(by = c("mode","Year"), nrow=1)+
  tmap_defaults +
  tm_layout(panel.label.height = c(0.8,0,0)) +
  tm_add_legend(type = "line", 
                col = c("grey40","lightblue"),
                labels = c("TAZ boundary","Community boundary"))+
  tm_add_legend(type = "symbol",
                col = "darkblue",
                labels = "TAZ centroid")
OD_flow_plot <- tmap_arrange(car_OD_flow_plot, walk_OD_flow, nrow = 2)
tmap_save(OD_flow_plot, paste0(here::here(),"/Figures/OD_flow_plot.png"), height=7, width=12, dpi=600)
```

```{r quartiles-of-tts-estimated-tt}
od_trips_perc_tt <- data.frame(x=c(gamma2011_motor$data, exp2011_nonmotor$data,
                                   gamma2016_motor$data, exp2016_nonmotor$data),
                               mode = c(rep("motorized", length(gamma2011_motor$data)),
                                        rep("non-motorized", length(exp2011_nonmotor$data)),
                                        rep("motorized", length(gamma2016_motor$data)),
                                        rep("non-motorized", length(exp2016_nonmotor$data))),
                               year = c(rep("2011", length(gamma2011_motor$data)),
                                        rep("2011", length(exp2011_nonmotor$data)),
                                        rep("2016", length(gamma2016_motor$data)),
                                        rep("2016", length(exp2016_nonmotor$data)))
)

q_motorized_2011 <- stats::quantile(od_trips_perc_tt |>
                                      filter(mode == "motorized", year == "2011") |>
                                      pull(x),
                                    probs = c(0.25, 0.5, 0.75), na.rm=T)

q_nonmotorized_2011 <- stats::quantile(od_trips_perc_tt  |>
                                         filter(mode == "non-motorized",
                                                year == "2011") |> 
                                         pull(x),
                                       probs = c(0.25, 0.5, 0.75), na.rm=T)

q_motorized_2016 <- stats::quantile(od_trips_perc_tt  |>
                                      filter(mode == "motorized",
                                             year == "2016") |>
                                      pull(x),
                                    probs = c(0.25, 0.5, 0.75), na.rm=T)

q_nonmotorized_2016 <- stats::quantile(od_trips_perc_tt |> 
                                         filter(mode == "non-motorized",
                                                year == "2016") |>
                                         pull(x),
                                       probs = c(0.25, 0.5, 0.75), na.rm=T)

q_1 <- data.frame(
  label = paste0("Q1:\n", c(q_motorized_2011[1], q_nonmotorized_2011[1]|> round(),
                            q_motorized_2016[1], q_nonmotorized_2016[1]|> round())),
  mode   = c("motorized", "non-motorized", "motorized", "non-motorized"),
  year = c("2011", "2011", "2016", "2016"),
  x     = c(q_motorized_2011[1], q_nonmotorized_2011[1] |> round(),
            q_motorized_2016[1], q_nonmotorized_2016[1]|> round()),
  y     = c(0.11, 0.11, 0.11, 0.11)
)

q_2 <- data.frame(
  label = paste0("Q2:\n", c(q_motorized_2011[2], q_nonmotorized_2011[2]|> round(),
                            q_motorized_2016[2], q_nonmotorized_2016[2]|> round())),
  mode   = c("motorized", "non-motorized", "motorized", "non-motorized"),
  year = c("2011", "2011", "2016", "2016"),
  x     = c(q_motorized_2011[2], q_nonmotorized_2011[2] |> round(),
            q_motorized_2016[2], q_nonmotorized_2016[2]|> round()),
  y     = c(0.11, 0.11, 0.11, 0.11)
)

q_3 <- data.frame(
  label = paste0("Q3:\n", c(q_motorized_2011[3], q_nonmotorized_2011[3]|> round(),
                            q_motorized_2016[3], q_nonmotorized_2016[3]|> round())),
  mode   = c("motorized", "non-motorized", "motorized", "non-motorized"),
  year = c("2011", "2011", "2016", "2016"),
  x     = c(q_motorized_2011[3], q_nonmotorized_2011[3] |> round(),
            q_motorized_2016[3], q_nonmotorized_2016[3]|> round()),
  y     = c(0.11, 0.11, 0.11, 0.11)
)
```

```{r tts-theoretical-curve-creation}
x <- data.frame(t = seq(0, 45, 0.1))
X_shorter <-data.frame(t = seq(0, 27, 0.1))

x_motorized_2011 <- x %>%
  mutate(f = dgamma(t,
                    gamma2011_motor$estimate[1],
                    gamma2011_motor$estimate[2]),
         mode = "motorized",
         year = "2011")

x_nonmotorized_2011 <- X_shorter %>%
  mutate(f = dexp(t,
                  exp2011_nonmotor$estimate[1]),
         mode = "non-motorized",
         year = "2011")

x_motorized_2016 <- x %>%
  mutate(f = dgamma(t,
                    gamma2016_motor$estimate[1],
                    gamma2016_motor$estimate[2]),
         mode = "motorized",
         year = "2016")

x_nonmotorized_2016 <- X_shorter %>%
  mutate(f = dexp(t,
                  exp2016_nonmotor$estimate[1]),
         mode = "non-motorized",
         year = "2016")

tld_theoretical <- rbind(x_motorized_2011,
                         x_nonmotorized_2011,
                         x_motorized_2016,
                         x_nonmotorized_2016) %>%
  mutate(mode = factor(mode, levels = c("motorized", "non-motorized")),
         year = factor(year, levels = c("2011", "2016")))
```

```{r ploting-empirical-and-theo-curve-TLD-tts, eval=FALSE}
TLD_empirical_wtheo_curve_plot <- ggplot() + 
  geom_histogram(data = od_trips_perc_tt,
                 aes(x = x, 
                     y = ..density..),
                 colour = "grey80",
                 alpha = 0.5,
                 boundary = 0,
                 binwidth = 3) +
  geom_vline(data = q_1, mapping = aes(xintercept = x), colour = "grey60") +
  geom_vline(data = q_2, mapping = aes(xintercept = x), colour = "grey60") +
  geom_vline(data = q_3, mapping = aes(xintercept = x), colour = "grey60") +
  geom_line(data = tld_theoretical, aes(x = t, y = f), colour = "blue", size=0.8) + 
  xlab("Estimated TTS travel time (minutes)") +
  scale_x_continuous(breaks = seq(0,40,10), expand=c(0,0))+
  scale_y_continuous(breaks = seq(0,0.12,0.04), limits = c(0,0.12))+
  facet_grid(~mode~year)+
  geom_text(data = q_1, 
            mapping = aes(x = x, y = y, label = label), colour = "grey20", size=3) +
  geom_text(data = q_2,
            mapping = aes(x = x, y = y, label = label), colour = "grey20", size=3) +
  geom_text(data = q_3,
            mapping = aes(x = x, y = y, label = label), colour = "grey20", size=3) +
  theme_classic()
TLD_empirical_wtheo_curve_plot
ggsave(paste0(here::here(),"/Figures/TLD_empirical_wtheo_curve_plot.png"), height=7, width=12, dpi=600)
```

```{r, eval=FALSE}
Vim_20112016_plot <-  tm_shape(SAvail_im_DA, bbox=SAvail_im_DA) +
  tm_polygons(col = "V_im", border.alpha = 0, border.col = "grey20", lwd=0.2,
              palette = "Greys",
              style="quantile", n=4,
              title = expression(V[i]^m*' (quartiles)'),
              textNA = "Unavailable",
              colorNA = "white")+
  tm_facets(by = c("mode","year"), nrow=1)+
  tm_shape(ALL_SCHOOLS_Elem_formapping) +
  tm_dots(col = "Status", size=0.03, title = "School status between\n2011 and 2016",
          palette = c("ivory", "black", "darkgreen","#0E86D4","#90EE90"),
          labels = c("No change","Closed before 2016", "Expanded before 2016","Moved before 2016","New after 2011"), legend.show=FALSE) +
  tm_shape(ham_community_bounds) + tm_borders(alpha = 0.8, col = "lightblue", lwd=1.5)+
  tmap_defaults +
  tm_layout(panel.label.height = 0.8)

vvim_20112016_plot <-  tm_shape(SAvail_im_DA, bbox=SAvail_im_DA) +
  tm_polygons(col = "v_im_kid_popadj", border.alpha = 0, border.col = "grey20", lwd=0.2,
              palette = "RdYlGn",
              midpoint = 1.0,
              title = expression(v[i]^m),
              breaks = c(0, 0.5, 1, 1.5, 2, 2.1), #note the highest value is ~12, I'm breaking the values early for the purpose of visualisation
              labels = c("[0-0.5)", "[0.5-1)", "[1-1.5)", "[1.5-2)", "[>2)"),
              textNA = "Unavailable",
              colorNA = "white")+
  tm_facets(by = c("mode","year"), nrow=1)+
  tm_shape(ALL_SCHOOLS_Elem_formapping) +
  tm_dots(col = "Status", size=0.03, title = "School status between\n2011 and 2016",
          palette = c("ivory", "black", "darkgreen","#0E86D4","#90EE90"),
          labels = c("No change","Closed before 2016", "Expanded before 2016","Moved before 2016","New after 2011")          ) +
  tm_shape(ham_community_bounds) + tm_borders(alpha = 0.8, col = "lightblue", lwd=1.5)+
  tmap_defaults +
  tm_layout(panel.label.height = 0.8)

Vim_vvim_20112016_plot <- tmap_arrange(Vim_20112016_plot, vvim_20112016_plot, nrow=2)

tmap_save(Vim_vvim_20112016_plot, paste0(here::here(),"/Figures/Vim_vvim_20112016_plot.png"), height=7, width=12, dpi=600)
```

```{r, eval=FALSE}
Vim_20112016_CHANGE_plot <-  tm_shape(SAvail_im_DA_changes, bbox=SAvail_im_DA) +
  tm_polygons(col = "c_V_im", border.alpha = 0, border.col = "grey20", lwd=0.2,
              palette = "-PuOr", midpoint=0,
              breaks = c(-5, -4, -3, -2, -1, 0, 1, 2,3,4,5),
              labels = c("[>-5 to -4)", "[-4 to -3)", "[-3 to -2)","[-2 to -1)","[-1 to 0)", "[0 to 1)", 
                         "[1 to 2)", "[2 to 3)", "[3 to 4)", "[4 to <5)" ),
              title = expression("Change in "*V[i]^m),
              textNA = "Unavailable",
              colorNA = "white")+
  tm_facets(by = c("mode"), nrow=1)+
  tm_shape(ALL_SCHOOLS_Elem_formapping |> 
             filter(Status != "NoChange" & Status != "Moved")|>
             mutate(Status_change = case_when(Status == "Removed" ~ "Closed",
                                              (Status == "Expanded" | Status == "New" ) ~ "New or Expanded"))) +
  tm_dots(col = "Status_change", size=0.05, title = "Schools that changed status",
          palette = c("black", "#90EE90"), legend.show=FALSE) +
  tm_shape(ham_community_bounds) + tm_borders(alpha = 0.8, col = "lightblue", lwd=2)+
  tmap_defaults +
  tm_layout(panel.label.height = 0.8)

#tmap_save(Vim_20112016_CHANGE_plot, "Figures/Vim_20112016_CHANGE_plot.png", height=7, width=12, dpi=600)

vvim_20112016_CHANGE_plot <- tm_shape(SAvail_im_DA_changes, bbox=SAvail_im_DA) +
  tm_polygons(col = "c_v_im_kid_popadj", border.alpha = 0, border.col = "grey20", lwd=0.2,
              palette = "RdYlGn", midpoint=0,
              breaks = c(-0.1, -0.05, 0.00, 0.05, 0.1),
              labels = c("[>-0.1 to -0.05)", "[-0.05 to 0.00)", "[0.00 to 0.05)", "[0.05 to <0.1)"),
              title = expression("Change in "*v[i]^m),
              textNA = "Unavailable",
              colorNA = "white")+
  tm_facets(by = c("mode"), nrow=1)+
  tm_shape(ALL_SCHOOLS_Elem_formapping |> 
             filter(Status != "NoChange" & Status != "Moved")|>
             mutate(Status_change = case_when(Status == "Removed" ~ "Closed",
                                              (Status == "Expanded" | Status == "New" ) ~ "New or Expanded"))) +
  tm_dots(col = "Status_change", size=0.05, title = "Schools that changed status",
          palette = c("black", "#90EE90")) +
  tm_shape(ham_community_bounds) + tm_borders(alpha = 0.8, col = "lightblue", lwd=2)+
  tmap_defaults +
  tm_layout(panel.label.height = 0.8)

Vim_vvim_20112016_CHANGE_plot <- tmap_arrange(Vim_20112016_CHANGE_plot, vvim_20112016_CHANGE_plot, nrow=2)

tmap_save(Vim_vvim_20112016_CHANGE_plot, 
          paste0(here::here(),"/Figures/Vim_vvim_20112016_CHANGE_plot.png"),
          height=7,
          width=12, 
          dpi=600)
```

```{r, eval=FALSE}
walk_OD_flow <- tm_shape(od_trips_perc_walk |> mutate(mode = "non-motorized"), bbox=SAvail_i_DA_extra) +
  tm_polygons(col = "TAZOrig_walk_perc", border.alpha = 0.4, border.col = "grey20", lwd=0.2,
              palette = "Greys",
              breaks = c(0, .25, .50, .75, 1),
              title = "Proportion of mode use",
              textNA = "Unavailable",
              colorNA = "white",
              legend.show=FALSE)+
  #tm_shape(SAvail_i_DA_extra) + tm_borders(alpha = 0.4, col = "cornflowerblue", lwd=0.2)+
  tm_shape(ham_taz) + tm_borders(alpha=0.4, col = "grey40", lwd=0.2) +
  tm_shape(ham_community_bounds) + tm_borders(alpha = 0.8, col = "lightblue", lwd=2)+
  tm_shape(ham_taz_centroids) + tm_dots(alpha=0.5, col="darkblue", size=0.01) +
  tm_shape(desire_lines_ham |> filter(TAZOrig_walk_trips >= 1) 
           |> mutate(mode = "non-motorized")) +
  tm_lines(palette = "Greens", breaks = c(12, 23, 44, 67, 174),
           title.col = "Number of trips (non-motorized)",
           col = "TAZOrig_walk_trips",
           alpha = 0.9,
           scale = 4,
           lwd = "all",
           legend.lwd.show = FALSE) +
  tm_facets(by = c("mode","Year"), nrow=1)+
  tmap_defaults +
  tm_layout(panel.label.height = c(0.8,0,0)) +
  tm_add_legend(type = "line", 
                col = c("grey40","lightblue"),
                labels = c("TAZ boundary","Community boundary"))+
  tm_add_legend(type = "symbol",
                col = "darkblue",
                labels = "TAZ centroid")
OD_flow_plot <- tmap_arrange(car_OD_flow_plot, walk_OD_flow, nrow = 2)
tmap_save(OD_flow_plot, 
          paste0(here::here(),"/Figures/SA-and-SAperStud-20112016.png"),
          height=7, 
          width=12,
          dpi=600)
```

```{r, eval=FALSE}
SA_Vim_2011_plot <- ggplot() +
  geom_sf(data = SAvail_im_DA2011, aes(fill= V_im), color = NA) +
  scale_fill_binned(type = "viridis", na.value = "grey60",
                    breaks = c(min(SAvail_im_DA2011$V_im),
                               quantile(SAvail_im_DA2011$V_im,0.25)%>%as.numeric(),
                               median(SAvail_im_DA2011$V_im, na.rm=T),
                               quantile(SAvail_im_DA2011$V_im,0.75)%>%as.numeric(),
                               max(SAvail_im_DA2011$V_im, na.rm = T)))+
  geom_sf(data = ALL_SCHOOLS_Elem |> filter(Year != 2016), 
          size=0.5, colour = "Red", alpha=0.7)+
  geom_sf(data = ham_community_bounds, fill=NA, color = "grey30", alpha=0.7)+
  theme_minimal() +
  theme(legend.title = element_text(size=10),
        axis.text = element_blank(), #removes lat/long from the axises
        panel.grid.major=element_blank())+
  facet_wrap(~mode)

SA_Vim_2016_plot <- ggplot() +
  geom_sf(data = SAvail_im_DA2016, aes(fill= V_im), color = NA) +
  scale_fill_binned(type = "viridis", na.value = "grey60",
                    breaks = c(min(SAvail_im_DA2016$V_im),
                               quantile(SAvail_im_DA2016$V_im,0.25)%>%as.numeric(),
                               median(SAvail_im_DA2016$V_im, na.rm=T),
                               quantile(SAvail_im_DA2016$V_im,0.75)%>%as.numeric(),
                               max(SAvail_im_DA2016$V_im, na.rm = T)))+
  geom_sf(data = ALL_SCHOOLS_Elem |> filter(Year != 2011), 
          size=0.5, colour = "Red", alpha=0.7)+
  geom_sf(data = ham_community_bounds, fill=NA, color = "grey30", alpha=0.7)+
  theme_minimal() +
  theme(legend.title = element_text(size=10),
        axis.text = element_blank(), #removes lat/long from the axises
        panel.grid.major=element_blank())+
  facet_wrap(~mode)
```
```{r, eval=FALSE}
SA_vim_2011_plot <- ggplot() +
  geom_sf(data = SAvail_im_DA2011, aes(fill= v_im_kid_popadj), color = NA) +
  scale_fill_binned(type = "viridis",
                    breaks = c(min(SAvail_im_DA2011$v_im_kid_popadj, na.rm = T),
                               quantile(SAvail_im_DA2011$v_im_kid_popadj,0.25)%>%as.numeric(),
                               median(SAvail_im_DA2011$v_im_kid_popadj, na.rm=T),
                               quantile(SAvail_im_DA2011$v_im_kid_popadj,0.75)%>%as.numeric(),
                               max(SAvail_im_DA2011$v_im_kid_popadj, na.rm = T)))+
  geom_sf(data = ALL_SCHOOLS_Elem |> filter(Year != 2016), 
          size=0.5, colour = "Red", alpha=0.7)+
  geom_sf(data = ham_community_bounds, fill=NA, color = "grey30", alpha=0.7)+
  theme_minimal() +
  theme(legend.title = element_text(size=10),
        axis.text = element_blank(), #removes lat/long from the axises
        panel.grid.major=element_blank())+
  facet_wrap(~mode)

SA_vim_2016_plot <- ggplot() +
  geom_sf(data = SAvail_im_DA2016, aes(fill= v_im_kid_popadj), color = NA) +
  scale_fill_binned(type = "viridis",
                    breaks = c(min(SAvail_im_DA2016$v_im_kid_popadj, na.rm = T),
                               quantile(SAvail_im_DA2016$v_im_kid_popadj,0.25)%>%as.numeric(),
                               median(SAvail_im_DA2016$v_im_kid_popadj, na.rm=T),
                               quantile(SAvail_im_DA2016$v_im_kid_popadj,0.75)%>%as.numeric(),
                               max(SAvail_im_DA2016$v_im_kid_popadj, na.rm = T)))+
  geom_sf(data = ALL_SCHOOLS_Elem |> filter(Year != 2011), 
          size=0.5, colour = "Red", alpha=0.7)+
  geom_sf(data = ham_community_bounds, fill=NA, color = "grey30", alpha=0.7)+
  theme_minimal() +
  theme(legend.title = element_text(size=10),
        axis.text = element_blank(), #removes lat/long from the axises
        panel.grid.major=element_blank())+
  facet_wrap(~mode)
```

```{r, eval=FALSE}
SA_vim_LIMAT_2011_plot <- ggplot() +
  geom_sf(data = SAvail_im_DA2011, aes(fill= v_im_kid_popadj_LIMAT), color = NA) +
  scale_fill_binned(type = "viridis", 
                    breaks = c(min(SAvail_im_DA2011$v_im_kid_popadj_LIMAT, na.rm = T),
                               quantile(SAvail_im_DA2011$v_im_kid_popadj_LIMAT,0.25)%>%as.numeric()+0.000000001,
                               median(SAvail_im_DA2011$v_im_kid_popadj_LIMAT, na.rm=T),
                               quantile(SAvail_im_DA2011$v_im_kid_popadj_LIMAT,0.75)%>%as.numeric(),
                               max(SAvail_im_DA2011$v_im_kid_popadj_LIMAT, na.rm = T)))+
  geom_sf(data = ALL_SCHOOLS_Elem |> filter(Year != 2016), 
          size=0.5, colour = "Red", alpha=0.7)+
  geom_sf(data = ham_community_bounds, fill=NA, color = "grey30", alpha=0.7)+
  theme_minimal() +
  theme(legend.title = element_text(size=10),
        axis.text = element_blank(), #removes lat/long from the axises
        panel.grid.major=element_blank())+
  facet_wrap(~mode)

SA_vim_LIMAT_2016_plot <- ggplot() +
  geom_sf(data = SAvail_im_DA2016, aes(fill= v_im_kid_popadj_LIMAT), color = NA) +
  scale_fill_binned(type = "viridis", 
                    breaks = c(min(SAvail_im_DA2016$v_im_kid_popadj_LIMAT, na.rm = T),
                               quantile(SAvail_im_DA2016$v_im_kid_popadj_LIMAT,0.25)%>%as.numeric()+0.000000001,
                               median(SAvail_im_DA2016$v_im_kid_popadj_LIMAT, na.rm=T),
                               quantile(SAvail_im_DA2016$v_im_kid_popadj_LIMAT,0.75)%>%as.numeric(),
                               max(SAvail_im_DA2016$v_im_kid_popadj_LIMAT, na.rm = T)))+
  geom_sf(data = ALL_SCHOOLS_Elem |> filter(Year != 2011), 
          size=0.5, colour = "Red", alpha=0.7)+
  geom_sf(data = ham_community_bounds, fill=NA, color = "grey30", alpha=0.7)+
  theme_minimal() +
  theme(legend.title = element_text(size=10),
        axis.text = element_blank(), #removes lat/long from the axises
        panel.grid.major=element_blank())+
  facet_wrap(~mode)

SA_Vim_LIMAT_2011_plot <- ggplot() +
  geom_sf(data = SAvail_im_DA2011, aes(fill= V_i_kid_popadj_LIMAT), color = NA) +
  scale_fill_binned(type = "viridis", 
                    breaks = c(min(SAvail_im_DA2011$V_i_kid_popadj_LIMAT, na.rm = T),
                               quantile(SAvail_im_DA2011$V_i_kid_popadj_LIMAT,0.25)%>%as.numeric()+0.000000001,
                               median(SAvail_im_DA2011$V_i_kid_popadj_LIMAT, na.rm=T),
                               quantile(SAvail_im_DA2011$V_i_kid_popadj_LIMAT,0.75)%>%as.numeric(),
                               max(SAvail_im_DA2011$V_i_kid_popadj_LIMAT, na.rm = T)))+
  geom_sf(data = ALL_SCHOOLS_Elem |> filter(Year != 2016), 
          size=0.5, colour = "Red", alpha=0.7)+
  geom_sf(data = ham_community_bounds, fill=NA, color = "grey30", alpha=0.7)+
  theme_minimal() +
  theme(legend.title = element_text(size=10),
        axis.text = element_blank(), #removes lat/long from the axises
        panel.grid.major=element_blank())+
  facet_wrap(~mode)

SA_Vim_LIMAT_2016_plot <- ggplot() +
  geom_sf(data = SAvail_im_DA2016, aes(fill= V_i_kid_popadj_LIMAT), color = NA) +
  scale_fill_binned(type = "viridis", 
                    breaks = c(min(SAvail_im_DA2016$V_i_kid_popadj_LIMAT, na.rm = T),
                               quantile(SAvail_im_DA2016$V_i_kid_popadj_LIMAT,0.25)%>%as.numeric()+0.000000001,
                               median(SAvail_im_DA2016$V_i_kid_popadj_LIMAT, na.rm=T),
                               quantile(SAvail_im_DA2016$V_i_kid_popadj_LIMAT,0.75)%>%as.numeric(),
                               max(SAvail_im_DA2016$V_i_kid_popadj_LIMAT, na.rm = T)))+
  geom_sf(data = ALL_SCHOOLS_Elem |> filter(Year != 2011), 
          size=0.5, colour = "Red", alpha=0.7)+
  geom_sf(data = ham_community_bounds, fill=NA, color = "grey30", alpha=0.7)+
  theme_minimal() +
  theme(legend.title = element_text(size=10),
        axis.text = element_blank(), #removes lat/long from the axises
        panel.grid.major=element_blank())+
  facet_wrap(~mode)
```
```{r, eval=FALSE}
SA_Vim_2011_plot
SA_vim_2011_plot
SA_Vim_LIMAT_2011_plot
SA_vim_LIMAT_2011_plot

SA_Vim_2016_plot
SA_vim_2016_plot
SA_Vim_LIMAT_2016_plot
SA_vim_LIMAT_2016_plot
#*summarize LIM at city level
```

```{r, eval=FALSE}
SAvail_im_DA2011  |> st_drop_geometry() |> group_by(COMMUNITY, mode) |> summarise(
  LIM = mean(DA_prev_LIMAT_under18, na.rm = T)*mean(kid_popadj_m, na.rm = T),
  sum_total_tt = sum(total_tt),
  emis_perkid = sum(emissions_total_car_tt)/sum(kid_popadj_m),
  sum_V_im = sum(V_im),
  mn_v_im = mean(v_im_kid_popadj),
  mn_v_im_LIM = mean(v_im_kid_popadj_LIMAT))

SAvail_im_DA2016  |> st_drop_geometry() |> group_by(COMMUNITY, mode) |> summarise(
  LIM = mean(DA_prev_LIMAT_under18, na.rm = T)*mean(kid_popadj_m, na.rm = T),
  sum_total_tt = sum(total_tt),
  emis_perkid = sum(emissions_total_car_tt)/sum(kid_popadj_m),
  sum_V_im = sum(V_im),
  mn_v_im = mean(v_im_kid_popadj),
  mn_v_im_LIM = mean(v_im_kid_popadj_LIMAT))                                                
```

```{r creating-desc-stats-table-tt, echo=FALSE}
#forming a complete descriptive statistic table
# Statistics <- data.frame("Statistic" = c("Parcel pts", "Trips",
#                                          "TT min.", "TT 1st qu.", "TT mean", "TT 3rd qu.", "TT max."))
# `2011` <- data.frame(`Y2011` = c(ALL_nearest_network_point %>% filter(Year != "2016") %>% pull(ROLL_NUM) %>% length() %>% round() ,
#                                 ALL_PARCELS %>% filter(ParYear != "2016" & Level == "Elementary") %>% pull(ROLL_NUM) %>% length()%>% round(),
#                                 ALL_PARCELS %>% filter(ParYear != "2016" & Level == "Elementary") %>% pull(car_tt) %>% min()%>% round(),
#                                 ALL_PARCELS %>% filter(ParYear != "2016" & Level == "Elementary") %>% pull(car_tt) %>% quantile(0.25) %>% as.numeric()%>% round(), 
#                                 ALL_PARCELS %>% filter(ParYear != "2016" & Level == "Elementary") %>% pull(car_tt) %>% mean()%>% round(),
#                                 ALL_PARCELS %>% filter(ParYear != "2016" & Level == "Elementary") %>% pull(car_tt) %>% quantile(0.75) %>% as.numeric()%>% round(),
#                                 ALL_PARCELS %>% filter(ParYear != "2016" & Level == "Elementary") %>% pull(car_tt) %>% max()%>% round()
#                                 ))
# `2016` <- data.frame(`Y2016` = c(ALL_nearest_network_point %>% filter(Year != "2011") %>% pull(ROLL_NUM) %>% length()%>% round(),
#                                 ALL_PARCELS %>% filter(ParYear != "2011" & Level == "Elementary") %>% pull(ROLL_NUM) %>% length()%>% round(),
#                                 ALL_PARCELS %>% filter(ParYear != "2011" & Level == "Elementary") %>% pull(car_tt) %>% min()%>% round(),
#                                 ALL_PARCELS %>% filter(ParYear != "2011" & Level == "Elementary") %>% pull(car_tt) %>% quantile(0.25) %>% as.numeric()%>% round(), 
#                                 ALL_PARCELS %>% filter(ParYear != "2011" & Level == "Elementary") %>% pull(car_tt) %>% mean()%>% round(),
#                                 ALL_PARCELS %>% filter(ParYear != "2011" & Level == "Elementary") %>% pull(car_tt) %>% quantile(0.75) %>% as.numeric()%>% round(),
#                                 ALL_PARCELS %>% filter(ParYear != "2011" & Level == "Elementary") %>% pull(car_tt) %>% max()%>% round()
#                                 ))
# desc_stats_tt <- cbind(Statistics,`2011`, `2016`)
# 
# #kable tabling
# desc_stats_tt %>%
#   kableExtra::kable(format = "latex",
#         align="lrrrrrr",
#         booktabs = T,
#         col.names = c("Statistic", "2011", "2016"),
#         caption = "\\label{tab:desc-stats-table-tt}Descriptive statistics of the parcels, trips, and calculated origin-destination car travel times.") %>%
#   kableExtra::kable_styling(full_width = "T",
#                 latex_options = c("scale_down"),
#                 position = "center")
```

# Introduction

Communities are by nature in flux: economic and demographic shifts, rural and urban configurations, and political pressures stoke changes in community characteristics. These forces spur the re-organization and consolidation of public facilities in communities around the world where changes to service provision may improve the quality of service for some but not for all [@Christiaanse_2020; @Rosik_Pulawska-Obiedowska_Goliszek_2021]. While change is natural, however, oftentimes top-down consolidation decisions that are driven by operational cost-savings do not take into consideration the full extent of losses in benefits.

From this perspective, school closure policies have been pursued by governments to optimize per-student operational cost-efficiency [@rongReviewResearchLowcarbon2022] but other implications, such as the impact of closures on property values [@MerrallWhatsASchool2024], the effects on the social infrastructure of communities [@butlerClosureRideau2019a; @irwinSchoolClosure2012], and particularly the travel implications of school closures have often been overlooked by decision makers [@bierbaumMobilityJusticeLinking2021; @Lee_Lubienski_2017]. This seems a mistake: journey-to-school trips are tremendously important, and travel for this purpose is the largest contributor of carbon emissions from personal travel after the journey-to-work [@rongReviewResearchLowcarbon2022; @pantelakiImpactHomeschoolCommuting2024]. In many communities globally, schools are closing [@sageman2022]. For instance, 65% of comprehensive schools (e.g. grades 1 through 9) in rural Finland since 1990 closed [@Autti_Hyry-Beihammer_2014], 10% of elementary schools (kindergarten through grade 8) in Chicago, USA in 2013 [@Lee_Lubienski_2017], and 7% of primary and lower secondary schools (kindergarten through grade 9) were closed in Denmark between 2010-2011 [@Beuchert_Humlum_Nielsen_Smith_2018]. School closures often increase the distance to school for many students and thus tend to induce motorized trips [@rongReviewResearchLowcarbon2022].

In this study, we focus on the transportation implications of school closure and consolidation policies. While some research has considered the "steps in reserve", that is, the potential for active travel hidden behind motorized trips [@morencyStepsReserve2009; @morencyHowMany2007], school closures potentially make actual steps vanish, sent to the reserve for the foreseeable future. In exchange for less physical activity, communities get more motorized travel, with the concomitant costs and burdens, including emissions. Quantification of losses in potential active travel and associated growith in carbon emissions is scarce in the literature, and the focus of this paper, which aims to make both methodological and empirical contributions. We take a novel multi-modal and singly-constrained accessibility measure, called spatial availability [@soukhovIntroducingSpatialAvailability2023; @soukhovMultimodalSpatialAvailability2024], and use it to estimate the changes in accessibility and travel before and after a recent wave of school closure/consolidations implemented in Hamilton, Ontario.

The results of the research indicate that, overall, school seats became less spatially available in the city. Though the school consolidation policy pursued by the province and implemente by the local school board may have resulted in cost-savings in operation and maintenance of facilities, families, students, and the environment now pay the price in terms of lost opportunities for healthier active travel, as well as increased emissions from motorized travel.

# Background

Hamilton is a mid-size city (\~570,000 pop) with rural, suburban and urban characteristics in the province of Ontario, Canada [@governmentofcanadaProfileTableCensus2022]. In 2013, the Hamilton-Wentworth District School Board (HWDSB),, the largest public English school board in the city, released its Long Term Facilities Master Plan [@hwdsbLongTermFacilities2013]. To ensure "equitable, affordable and sustainable learning facilities", the Master Plan indicated that 80% of it's elementary schools (i.e., schools attended by students aged 5 to 13) would be subjected to accommodation evaluations, including an "Accommodation Review" whose outcome historically has led to school closures/consolidation [@hwdsbLongTermFacilities2013; @craggsBoardEvaluatingFuture2013; @seasonsSchoolClosurePolicy2014]. These decisions by the school board were in part motivated by a search for operational savings based on projected reductions in student population and under-utilized school capacity [@craggsAxeCouldFall2012]. However, the case of Hamilton and the HWDSB was unlike others in the province: the HWDSB underwent an unprecedented number of Accommodation Reviews relative to other school boards [@seasons2014ARCCatalogues2014].

Though the Accommodation Review guidelines outline a community-focused process to deliberate a school's future viability to students, the community and the economy [@seasonsSchoolClosurePolicy2014; @ministryofeducationPupilAccommodationReview2006], concerned residents felt a wholesome breakdown of costs to repair schools was not made consistently available [@kleinhuisSeriousConcernsSchool2013]. Further, others pointed to flawed incentive structures motivating the HWDSB: "By closing three schools the HWDSB would have a strong case with the Ministry of Education to receive full funding for a new school" [@kleinhuisSeriousConcernsSchool2013]. Along with insufficient funding to maintain schools in a state of good repair [@auditorgeneralofontarioAnnualReport20152015], the operational savings as a case for school closures is a long-standing critique of the school "funding formula" established by the conservative provincial government in 1998 [@mackenzieBlueprintFixOntario2018; @irwinSchoolClosure2012]. With the backdrop of budget cuts, the funding formula pits financing of school maintenance and student needs against the financial compensation of staff and powers of local school boards as they depend on the province for the allocation of resources [@mackenzieBlueprintFixOntario2018].

In the wave of school accommodation reviews between 2013 and 2016, `r ALL_SCHOOLS_Elem %>% filter(Year == "2011" & Level == "Elementary" & Status == "Removed") %>% pull(SchoolID) %>% length()` elementary schools closed. These closures represented a `r ((ALL_SCHOOLS_Elem %>% filter(Year != "2016") %>% pull(SchoolID) %>% length() - ALL_SCHOOLS_Elem %>% filter(Year != "2011") %>% pull(SchoolID) %>% length())/ALL_SCHOOLS_Elem %>% filter(Year != "2016" & Level == "Elementary") %>% pull(SchoolID) %>% length()) %>% scales::percent()` and `r  (( ((avail_2011 |> group_by(SchoolID) |> summarize(OTGC = first(OTGC2011)) |> select(OTGC) |> sum())) - (avail_2016 |> group_by(SchoolID) |> summarize(OTGC = first(OTGC2016)) |> select(OTGC) |> sum())) / (avail_2016 |> group_by(SchoolID) |> summarize(OTGC = first(OTGC2016)) |> select(OTGC) |> sum()) ) |> scales::percent()` decline in elementary school locations and capacity between 2011 (before the school closures) and 2016, despite the fact that the overall student population in the city did not decline, and in fact increased slightly between 2011 (`r avail_2011$kid_popadj %>% sum(na.rm=T) %>% prettyNum(big.mark=",")` students) and 2016 (`r avail_2016$kid_popadj %>% sum(na.rm=T) %>% prettyNum(big.mark=",")` students).

While accommodation reviews were undertaken to reduce operational costs, how might the closure of schools impact local communities? In this paper, we hypothesize that the school consolidation policy reduced accessibility to schools for many students. By extension, this loss in access and the consequent displacement of active trips to school increased transportation-related emissions. To test these hypotheses, the paper examines the following: how does the (1) spatial availability, (2) loss in walk minutes and (3) associated carbon emissions change for home-to-school trips after a city closes and consolidates schools? Further, this work examines (4) what communities, based on low-income prevalence, experienced the greatest losses as a result of this policy.

Spatial availability, a singly-constrained measure of spatial accessibility [@soukhovIntroducingSpatialAvailability2023; @soukhovMultimodalSpatialAvailability2024], is used to quantify accessibility before- and after- school closures, and the associated impacts for the city of Hamilton and the elementary school-aged student population as a result of a policy with narrowly defined benefits. The methods used in this paper can be applied by decision-makers to evaluate the spatial and transport-related impacts of public service provision policy. Furthermore, to increase the diffusion of this analysis, all work is transparent and reproducible, following best practices in the spatial sciences [@brunsdonOpeningPractice2021; @paezOpenSpatial2021c]. Accordingly, all associated code and data will be publicly available within the lead author's [GitHub repository](https://github.com/soukhova)

# Data

Our focus is to evaluate the impact of school consolidations between 2011 and 2016 on multi-modal school spatial availability and travel-related GHG emission in Hamilton. To facilitate comparison, two scenarios are generated. Each scenario reflects the conditions in 2011 and 2016 respectively, assuming the school configurations in each year, as well as average student population and low-income prevalence of households, residential parcel locations, and estimated modal travel times. The difference between the two scenarios is quantified in terms of accessibility, emissions and equity implications and discussed in context with the school consolidation policy. Data were retrieved at different levels of spatial granularity from the following sources.

## Schools

First, schools and associated school catchments were retrieved from the City's school boards. In Hamilton, the majority of student-aged population attends a school in one of the two English public school boards. One board is referred to as *public* (i.e., Hamilton Wentworth District School Board (HWDSB)) in our work and the other is referred to as *public-catholic* (Hamilton Wentworth Catholic District School Board (HWCDSB)). These catchments indicate what residential property is assigned to what school, by default. Families can decide to attend schools out of catchment; in fact, 21%-23% <!--calculated in 03-travel-times-TTS.qmd--> of motorized school trips are out of catchment according to the regional travel survey, the Transportation Tomorrow Survey (TTS) in 2011 and 2016 [@data_management_group_tts_2018]. The elementary school locations and catchments were provided by the HWDSB and HWCDSB for the 2010-2011 and 2015-2016 academic years, and are shown in @fig-Fig1.

Formally, elementary schools are defined as schools that provide instruction to any combination of grades between kindergarten to grade 8 (i.e. typically children aged 5 to 14). As such, elementary includes middle schools that only instruct grades 6 to 8 (*Mid*), primary schools that only instruct kindergarten to grade 5 or 6 (*JrElem*), and all grade elementary schools (*Elem*) which instruct all grades from kindergarten to grade 8. The maximum number of students that may be enrolled in a school is calculated by the provincial Ministry of Education; this value is referred to as the on-the-ground capacity (*OTGC*) and is assigned to each school <!-- [AP]: what are the units of OTGC? Students? -->. The current/historic OTGCs are not publicly available [@ontarioSchoolFacilityInventory2017]. However, for schools that underwent the full Accommodation Review process, OTGCs for a given year could be obtained from publicly available School Information Profiles. A regression model was used to estimate the OTGCs for schools that lacked capacity information. The independent variables in the model were 1) the school's building footprint (*F*) in units of ($m^2$); 2) the school's shortest Euclidean distance from the centroid of Hamilton's central business district area (*DistCBD*) defined in units of (m); and 3) coefficients associated with a binary variable indicating the type of school grade instruction (*Mid*, *JrElem*, *Elem*). The building footprint from an archived spatial data set [@DMTI_Spatial_2015] and the 2016 footprints were retrieved from OSM [@OpenStreetMap_2021]. Because the age of construction for different schools is not available, we used each school's distance from the centroid of the Downtown Business Improvement Area [@Hamilton_2014] as a proxy for school age as the oldest buildings are generally located closer to the city centre [@Merrall_2021].

Equations (\ref{eq:OTGC-public}) and (\ref{eq:OTGC-catholic}) describe the estimated OTGC for public schools and catholic schools respectively. School status and OTGC summed by catchment is also visualised in @fig-Fig1. As seen in Table <!--\ref{TabA1-OTGC}--> in the Appendix, the coefficient of determination for these models was $R^2 = $ and $R^2 = $ for the Public and Catholic Boards respectively, and the standard deviation was $\sigma = $ and $sigma = $.

<!-- [AS]: shall we explain further how we valuated this model? I don't think its entirely necessary. But the code is here: https://github.com/soukhova/hamONdests/blob/master/data-raw/01-school-data-preparation.Rmd (Antonio is a collaborator) -->

<!-- [AP]: Just summarize the results here: what was the R2 and standard deviations of these models? -->

```{=tex}
\begin{equation}
\label{eq:OTGC-public}
OTGC_{Public} = F^{0.346}-e^{0.00003*DistCBD}+e^{3.123*JrElem_b}+e^{3.752*Elem_b}+ e^{3.068*Mid_b}
\end{equation}
```

```{=tex}
\begin{equation}
\label{eq:OTGC-catholic}
OTGC_{Catholic} =F^{0.471}-e^{0.00003*DistCBD}+e^{2.333*Elem_b}
\end{equation}
```

```{r}
#| label: fig-Fig1
#| fig-cap: "The school catchments, estimated on the ground capacities (OTGCs) of schools summed per catchment, and school status for both public and public catholic school boards for year 2011 and 2016. OTGCs is presented by quartiles."
#| echo: false
#| out-width: "140%"

knitr::include_graphics(paste0(here::here(),"/Figures/catchment_schools_plot.png"))
```
<!--perhaps in the legend of this map, I add the number of schools in brackets (for "school status between 2011 and 2016")? And the number of schools on the map in each box? what do you think? I'll also make the font on the panels bigger-->

<!-- Number of schools? Yes Also, make the symbols that map the schools bigger, and use different shapes in addition to different colors. As an alternative, the OTGC could be the color, and the school category the shape. -->

## Students, residential locations and low-income prevelance

In addition to OTGC for schools, statistics about the student population and proportion of low-income after tax (LIM-AT) housholds were retrieved from the 2011 and 2016 Canadian Census [@Statistics_Canada_2011; @Statistics_Canada_2016] using the {cancensus} `R` package [@von_Bergmann_Shkolnik_Jacobs_2021]. The census releases population data by age group category in 5-year increments. The LIM-AT prevalence for households with children under 18 (there is no LIM-AT prevalence tabulated for households with exclusively elementary aged children) is retrained. LIM-AT is refers to the proportion of private households that are below the median after-tax income [@governmentofcanadaDictionaryCensusPopulation2017]. These census variables are taken at the Dissemination Area (DA) level, which is the finest geographic unit for which census data are made publicly available. DAs are designed by Statistics Canada with the aim of population uniformity hence DAs greatly vary in area but represent between approximately 400 and 700 (1st and 3rd quartile) in total population. The population aged 5 to 14 and the proportion of LIM-AT are visualised in the first and last row in @fig-Fig2.

The third source of data consists of centroids of residential parcels in 2011 and 2016 that are used to represent residential locations [@Teranet_2009]. There are 134,340 and 139,467 residential locations in 2011 and 2016 respectively, representing `r cen_data_2011_hamiltononly$DA_households |> sum() |> prettyNum(big.mark = ",")` and `r cen_data_2016_hamiltononly$DA_households |> sum() |> prettyNum(big.mark = ",")` dwellings (according to the Canadian Census [@Statistics_Canada_2011; @Statistics_Canada_2016]). As the number of children per parcel are publicly unavailable, the average number of 5 to 14 year olds for each DA is divided by the number of residential households in that DA. Then the potential number of children at each dwelling is calculated by multiplying this DA rate by the number of dwellings in each parcel. <!-- NOTE: We consider the parcel centroids instead of DA centroids to more precisely calc. travel times. Since we do not know parcel student population, we assume a uniform spatial distribution of population across each DA. This assumption is a limitation, but it is better than what's typically done in studies that only consider the centroid of the census track. --> Due to the proprietary nature of the parcel data, the middle row in @fig-Fig2 visualises an aggregation of the information: the average rate of 5-14 year olds per residential household at the DA level.

```{r}
#| label: fig-Fig2
#| fig-cap: "The magnitude (top row) and rate per residential unit (middle row) of elementary aged student population per DA in 2011 and 2016. The proportion of low-income households with children under 18 years old is also shown (bottom row). All scales are represented in quartiles."
#| echo: false
#| out-width: "140%"

knitr::include_graphics(paste0(here::here(),"/Figures/stud_pop_rate_LIM_plot.png"))
```

## Estimated travel times and emissions

It is assumed that children can go to any elementary school, but there is a preference for facilities that are closer. Based on this assumption, the fourth source of data is information regarding the mode-used and origin-destination locations of home-to-school trips from the 2011 and 2016 TTS [@data_management_group_tts_2018]. The TTS is a travel survey conducted in the Greater Golden Horseshoe Area in Ontario every 5 years. With a target 5% sampling rate, the survey is expanded to be representative at the traffic analysis zone (TAZ) level of geography. TAZ are spatial units created for the purpose of the TTS; a few DAs typically nest within each TAZ. The trip-level travel data extract for this paper represent `r od_trips_perc_2011$TAZOrig_car_trips |> sum() |> prettyNum(big.mark = ",")` and `r od_trips_perc_2016$TAZOrig_car_trips |> sum() |> prettyNum(big.mark = ",")` motorized trips (mode used includes private care passenger, school bus, taxi, and transit) and `r od_trips_perc_2011$TAZOrig_walk_trips |> sum() |> prettyNum(big.mark = ",")` and `r od_trips_perc_2016$TAZOrig_walk_trips |> sum() |> prettyNum(big.mark = ",")` non-motorized trips (mode used includes walk and cycling) for 5 to 13 year olds from home to school in 2011 and 2016 respectively. The travel times for motorized and non-motorized travel are estimated using the {r5r} [@Pereira2021r5r] package, which interfaces with the java-based R5 routing engine developed separately by Conveyal [@conveyalConveyalR5Routing2022]. Motorized and non-motorized travel times are estimated for all TAZ centroids to all TAZ centroids and these estimated travel times are tied to the associated OD trips. The travel times and the intensity of modal home-to-school flows are visualised in @fig-Fig3 along with the boundaries of the TAZ and DAs. As can be seen in @fig-Fig3, not all TAZs capture an elementary school trip by both modes. For this reason, the proportion of modal-split is aggregated at the community level. Hamilton (Ancaster `r community_car_perc_2011$TAZOrig_car_perc[1] |> percent()`-`r community_car_perc_2016$TAZOrig_car_perc[1] |> percent()`, Dundas `r community_car_perc_2011$TAZOrig_car_perc[2] |> percent()`-`r community_car_perc_2016$TAZOrig_car_perc[2] |> percent()`, Flamborough `r community_car_perc_2011$TAZOrig_car_perc[3] |> percent()`-`r community_car_perc_2016$TAZOrig_car_perc[3] |> percent()`, Glanbrook `r community_car_perc_2011$TAZOrig_car_perc[4] |> percent()`-`r community_car_perc_2016$TAZOrig_car_perc[4] |> percent()`, Hamilton Central `r community_car_perc_2011$TAZOrig_car_perc[5] |> percent()`-`r community_car_perc_2016$TAZOrig_car_perc[5] |> percent()`, and Stoney Creek `r community_car_perc_2011$TAZOrig_car_perc[6] |> percent()`-`r community_car_perc_2016$TAZOrig_car_perc[6] |> percent()`). <!--Do we maybe need to discuss how this lack of data for each TAZ impacted the calculations? [AP]: No. Unrelated: did you use information from TTS2016R? Or can you add it to the data package? And then cite your data paper.-->

```{r}
#| label: fig-Fig3
#| fig-cap: "The origin-destination flows from home-based motorized and non-motorized school trips for students between 5-13 years old as retrieved from TTS 2011 and 2016 mapped atop proportion of TAZ non/motorized modal share, DA boundaries, and community boundaries."
#| echo: false
#| out-width: "140%"

knitr::include_graphics(paste0(here::here(),"/Figures/OD_flow_plot.png"))
```

The fifth source of data consists of estimated travel time matrices from all residential parcel locations to all elementary schools for both motorized (car) and non-motorized (walking) travel. The travel times are estimated using the {r5r} package [@Pereira2021r5r]. For both modes, the inputs assume a maximum travel time threshold of 60 minutes and the free-flow OpenStreetMap road network of Hamilton (retrieved using Geofabrik [@geofabrikOntarioOpenStreetMapGeofabrik2022] and manually edited to only include the Hamilton boundary). Further, only walking trips shorter than 27 min minutes were retained. This travel time is approximately equivalent to 1.6 km which is the point at which within-catchment students qualify for transport provided by the school board [@HWDSB_2019]. Estimated trip-weighted motorized travel times are on average 16.5 minutes in 2011 and 16.8 minutes in 2016 (summary statistics for both years are approximately min: 0.5, Q1: 11.0, Q2: 16.0, Q3: 21.0, max: 60.0). Estimated trip-weighted non-motorized travel times are on average 18.0 minutes in 2011 and 18.1 minutes in 2016 (summary statistics for both years are approximately min: 0.5, Q1: 13.0, Q2: 19.0, Q3: 24.0, max: 27.0).
<!--to calc. the summary statistics i loaded "OD_2011" and "OD_2016" from data-products (big files) and ran: quantile(OD_2011|>filter(mode=="w" & travel_time_p50 <= 27)|>pull(travel_time_p50), weights=OD_2011|>filter(mode=="w")|>pull(kid_popadj)) 
and 
mean(OD_2011|>filter(mode=="w" & travel_time_p50 <= 27)|>pull(travel_time_p50), weights=OD_2011|>filter(mode=="w")|>pull(kid_popadj))
-->

These travel time matrices along with the origin-destination flows were used to create modal trip length distribution (TLD) functions. Analysis of the TLD is a useful technique to calibrate impedance functions as they are a representation of the propensity of travel by trip length [@horbachov_theoretical_2018; @batista_estimation_2019]. The empirical TLD is used to fit a density distribution using maximum likelihood and moment matching techniques and the Nelder-Mead and Brent methods for direct optimization available within the {fitdistrplus} `R` package [@fitdistrplus_2015]. Based on goodness-of-fit criteria and diagnostics <!-- should I include the diagonostic plots in the appendix? q-q, etc..? I don't think its entirely necessary as the empirical is graphed over the theoretical line in the following figure. [AP] Agreed-->, the gamma and exponential distributions were selected for the motorized and non-motorized modal distributions respectively. The gamma distribution is defined by the shape ($\alpha$) parameter of `r round(gamma2011_motor$estimate[1], 3)` (2011) and `r round(gamma2016_motor$estimate[1], 3)` (2016) and the rate ($\beta$) of `r round(gamma2011_motor$estimate[2], 3)` (2011) and `r round(gamma2016_motor$estimate[2], 3)` (2016). The exponential distribution is defined by the rate ($\beta$) parameter of `r round(exp2011_nonmotor$estimate[1], 3)` (2011) and `r round(exp2016_nonmotor$estimate[1], 3)` (2016). The TLDs for the motorized and non-motorized trips are shown in @fig-Fig4. For reference, the gamma distribution and the exponential distribution function are displayed in the following Equation (\ref{eq:exp-dist})) and (\ref{eq:gamma-dist})) where $x$ is $c_{ij}$:

```{=tex}
\begin{equation}
\label{eq:exp-dist}
f(x) = \beta e ^{-\beta x}
\end{equation}
```

```{=tex}
\begin{equation}
\label{eq:gamma-dist}
\begin{aligned}
f(x, \alpha, \beta) = \frac {x^{\alpha-1}e^{-\frac{x}{\beta}}}{ \beta^{\alpha}\Gamma(\alpha)} \quad \text{for } 0 \leq x \leq \infty \\
\Gamma(\alpha) =  \int_{0}^{\infty} x^{\alpha-1}e^{-x} \,dx
\end{aligned}
\end{equation}
```

```{r}
#| label: fig-Fig4
#| fig-cap: "Empirical (grey bars) and theoretical (blue curves) motorized and non-motorized impedance functions. Motorized theoretical impedance function is based on the gamma distribution function and non-motorized on the expoential distribution function"
#| echo: false
#| out-width: "140%"

knitr::include_graphics(paste0(here::here(),"/Figures/TLD_empirical_wtheo_curve_plot.png"))
```
<!--maybe bigger font here?-->

The sixth and final data source relates to motorized GHG well-to-wheel estimations. It assumed that all non-motorized trips produce no emissions and those that are motorized take the average emission factor of 0.1 kg CO2e per minute travel. This emission factor is assumed based on an average speed of 40 km/h, the 2,380 g $CO_{2}e$ produced when 1 L of gasoline is combusted, and an internal combustion engine passenger vehicle fuel efficiency of 7 g/100km as discussed in @soukhovOccupancyGHGEmissions2022.

# Methods: Multi-modal spatial availability

Accessibility is defined as the "potential for spatial interaction". It is classically presented as the gravity-based measure popularized by @hansenHowAccessibilityShapes1959 and takes the following general multi-modal formulation (Equation (\ref{eq:multi-modal-conventional-accessibility})):

```{=tex}
\begin{equation}
\label{eq:multi-modal-conventional-accessibility}
S_i^m = \sum_{j=1}^JO_j \cdot f^m(c_{ij}^m)
\end{equation}
```
\noindent where:

-   $m$ is a set of modes.
-   $c_{ij}^m$ is a measure of the cost of moving between $i$ and $j$ for each $m$.
-   $f^m(\cdot)$ is an impedance function of $c_{ij}^m$ for each $m$; it can take the form of any monotonically decreasing function chosen based on positive or normative criteria [@paez2012measuring].
-   $i$ is a set of origin locations ($i = 1,\cdots,N$).
-   $j$ is a set of destination locations ($j = 1,\cdots,J$).
-   $O_j$ is the number of opportunities at location $j$.
-   $S$ is Hansen-type accessibility as weighted sum of opportunities.

As indicators of urban structure, Hansen-type measures are informative, but the meaning of 10,000 accessible school-seats is harder to pin down: how many opportunities must any single student have access to? Furthermore, in scenario comparisons, the interpretation of the percentage change in accessible school seats is unclear. The interpretability of Hansen-type accessibility has been discussed in numerous studies, including recently by @hu_2019_measuring, @kelobonye2020measuring, and in greater depth by @merlin2017competition along with @soukhovIntroducingSpatialAvailability2023 and @soukhovMultimodalSpatialAvailability2024. The interpretation of accessibility is suggested to depend on how many people demand the opportunity, especially for exclusive opportunity-types like schools-seats (i.e., one school-seat is for one student).

In this paper, we benefit from new developments in accessibility research, particularly the multi-modal spatial availability measure [@soukhovIntroducingSpatialAvailability2023; @soukhovMultimodalSpatialAvailability2024]. Spatial availability is a singly-constrained accessibility measure that accounts for competition by students using different modes for exclusive opportunities, such as school seats. The measure’s single constraint ensures that the marginals at the destination are met and thus the number of estimated school seats (opportunities) are preserved and allocated proportionally to the mode-using student population. This proportional allocation of opportunities yields an interpretable and meaningful measure of opportunity access, particularly when comparing across modes, at the spatial resolution of a residential parcel, and multiple time periods. See @soukhovMultimodalSpatialAvailability2024 for further discussion, multi-modal spatial availability $V_{i}^m$ is defined as given by Equation (\ref{eq:spatial-availability}).

```{=tex}
\begin{equation}
\label{eq:spatial-availability}
V_{i}^{m} = \sum_{j=1}^J O_j\ F^{tm}_{ij}
\end{equation}
```
\noindent where:

-   $F^{tm}_{ij}$ is a balancing factor that depends on the population and cost of movement in the system as part of the gravity modelling framework and is captured in Equation (\ref{eq:balancing-factors}) for mode $m$.
-   $O_j$ is the number of opportunities at $j$.
-   $V_i^m$ is the number of spatially available opportunities from the perspective of $i$ for mode $m$.

$F^{tm}_{ij}$ can be understood as the joint probability of allocating opportunities, where $F^{pm}_{i}$ is the population-based balancing factor that grants a larger share of opportunities to larger $m$ population spatial units and $F^{cm}_{ij}$ is the impedance-based balancing factor that grants a larger share of the opportunities to less $m$-travel costly centers. Together $F^{tm}_{ij}$ ensures proportional allocation such that opportunities $O$ (like school-seats) are preserved for the whole region (i.e., $O = \sum_{j} O_j = \sum_{i} V_i = \sum_{m}\sum_{i} V_i^m$ ) and is reflected in Equation (\ref{eq:balancing-factors}):

```{=tex}
\begin{equation}
\label{eq:balancing-factors}
F^{tm}_{ij} = \frac{F^{pm}_{i} \cdot F^{cm}_{ij}}{\sum_{i} F^{pm}_{i} \cdot F^{cm}_{ij}}
\end{equation}
```
\noindent where:

- The factor for allocation by population for each $m$ at each $i$ is $F^{pm}_{i} = \frac{P_{i}^m}{\sum_{m}\sum_{i} P_{i}^m}$
- The factor for allocation by travel cost for each $m$ at each $i$ and $j$ is $F^{cm}_{ij} = \frac{f^m(c^m_{ij})}{\sum_{m}\sum_{i} f^m(c^m_{ij})}$

It should be noted that, when summed over all spatial units in the region, the population-based allocation factors $F^{pm}_{i}$ always equal 1 (i.e., $\sum_{m}\sum_{i} F^{pm}_{i}= 1$), likewise for impedance-based allocation factors $F^{cm}_{i}$ (i.e., $\sum_{m}\sum_{i} F^{cm}_{ij} = 1$).

Hansen-type accessibility is not designed to preserve the number of opportunities in the region, it simply counts the intensity of opportunities that those in a zone can potentially interact with (weighted by the friction of distance). Also, as discussed in @soukhovIntroducingSpatialAvailability2023, popular competitive accessibility measures such as the 2SFCA [@shenLocationCharacteristicsInnercity1998; @luoMeasuresSpatialAccessibility2003] are internally inconsistent, and the only way it preserves the number of opportunities is if the effect of the impedance function is ignored when expanding the values of opportunities per capita to obtain the total number of opportunities. On the other hand, the proportional allocation procedure associated with calculating multi-modal spatial availability $V_i^m$ consistently returns a number of opportunities available to populations by mode that matches the total number of opportunities in the region when summed. By doing this consistently, it is possible to define a measure of multi-modal spatial availability per capita ($P$) as presented in Equation (\ref{eq:SA-per-capita}) for use as a benchmark to compare against the regional opportunities per capita ($\frac{\sum_{j} O_j}{\sum_{i} P_i}$).

```{=tex}
\begin{equation}
\label{eq:SA-per-capita}
v_i^m = \frac{V_i^m}{P_i^m}
\end{equation}
```

This paper uses the multi-modal spatial availability $V_i^m$ and spatial availability per student $v_i^m$ measures to quantify accessibility changes for the following reasons. First, all residential locations are associated to their respective census DA, TTS TAZ, and community boundary (@fig-Fig3) based on spatial location. Each parcel is assigned a number of 'potential' motorized and non-motorized student aged population from these different data sources to calculate a motorized and non-motorized population balancing factor $F_{i}^{pm}$. For each parcel the following is retained: the 2011 and 2016 census average elementary-aged student population per parcel (@fig-Fig2 second row) and the average motorized and non-motorized split as informed by the TTS aggregated by community (@fig-Fig3). Second, the school capacity $O_j$ is estimated and associated with each school (@fig-Fig1). All residential locations are assumed to be able to access all schools by either motorized or non-motorized mode. All origins (residential locations) can reach all schools by motorized mode, but few residential locations can reach schools within a 27 minute non-motorized trip.

Third, the motorized and non-motorized impedance-balancing factors $F_{ij}^{cm}$ are calculated using a mode-specific impedance function (@fig-Fig4) based on the respective travel times for each origin (parcel) to school combination. Finally, outputs from all three stages are joined together. $F_{i}^{pm}$ joined with $F_{ij}^{cm}$ yields the total balancing factor $F^{tm}_{ij}$ which serves to proportionally allocate the school capacity $O_j$ to each parcel. The value is multi-modal spatial availability $V_i^m$ which can be interpreted as the number of school-seats that are available to the mode-using population at that parcel. $V_i^m$ is then summed and represented at the DA and community level along with the calculated $v_i^m$ for interpretation.

# Results

```{r}
#| label: fig-Fig6
#| fig-cap: "Multi-modal spatial availability (greys) and spatial availability per mode-using student (diverging reds and greens) aggregated at the level of DA for 2011 and 2016."
#| echo: false
#| out-width: "140%"
#| 
knitr::include_graphics(paste0(here::here(),"/Figures/Vim_vvim_20112016_plot.png"))
```

<!--I'll make this plot longer-->

The motorized and non-motorized spatial availabilities of school seats $V_i$ for both 2011 and 2016 are presented in the top half of @fig-Fig6. Though they show different magnitudes, both populations using non- and motorized modes have higher spatial availability when they are more proximate to schools. Hence, all DAs have higher values within Hamilton Central and those more proximate to schools in less-urban communities. Trends appear similar for both 2011 and 2016: proximate populations using motorized modes benefit from the ability to reach more distant destinations than non-motorized modes. When $V_i^m$ is divided by the number of students by mode in that DA, $v_i^m$ is derived (bottom half of @fig-Fig6). A similar spatial distribution seen in $V_i^m$ can be seen in $v_i^m$ plots, but as a rate of student population. If our aim is to provide sufficient availability to schools for both motorized and non-motorized populations (i.e., greater than 1.0 $v_i^m$ school-seat availability per student), this rate is only achieved in the core of the city for motorized modes, and notably only for non-motorized populations in DAs that are in less-densely populated rural areas that are school proximate and certain pockets of Hamilton Central. The majority of schools closed were within Hamilton Central, so non-motorized populations in DAs within Hamilton Central were the most negatively impacted by losses in availability while certain motorized populations captured these losses as availability gains.

```{r}
#| label: fig-Fig7
#| fig-cap: "Change in multi-modal spatial availability and spatial availability per mode-using student between 2016 and 2011 at the level of DA of the 2016 Canadian Census."
#| echo: false

knitr::include_graphics(paste0(here::here(),"/Figures/Vim_vvim_20112016_CHANGE_plot.png"))
```

<!--The colours combos I choose here are absolutely ugly ;( but besides that, do you agree that V_im and v_im should use different colour palettes?-->

To emphasize how spatial availability changed between 2016 and 2011, the change in $V_i^m$ and $v_i^m$ is visualised for both modes in @fig-Fig7. Overall, a decrease in both $V_i^m$ and $v_i^m$ can be seen across both modes around closed schools (black points). An increase near new or expanded schools (green points) is not as notable as overall fewer school seats were added relative to increases in the school-aged population, leading to increased competition for these resources. For non-motorized mode-using populations, $V_i^m$ decreases across the city. However, some motorized-mode using populations in DAs do experience an increase in $V_i^m$: these DAs are often relatively proximate to schools but more immediately outside walking-range. Under competitive and constrained calculations, these DAs benefit from the losses in spatial availability that are predominately experienced by the non-motorized populations within Hamilton Central.

<!--Antonio, what do you think, maybe worth adding a plot of GHG emissions or change in tt? This will provide us some higher level summary stats. Moataz asked about this as well regarding the figures in the Results section. [AP] Yes, please add those statistics to support the discussion. -->

<!--Antonio, what do you think, add a table with summaries of LIMAT, Kid pop, V_im change, v_im change, total travel time change, and emission changed by community and mode? We should discuss to decide what's the best presentation. [AP]: Best presentation: spatial availability by quartiles of the relevant dimensions (e.g., LIMAT) -->

```{r}
# SAvail_im_COMM_changes |> filter(mode == "motorized") |> st_drop_geometry() |> 
#   dplyr::select(c("COMMUNITY",
#                   "prev_LIMAT_2016", 
#                   "c_kid_perc", 
#                   "c_V_im_perc", 
#                   "c_v_im_perc", 
#                   "emissions_2011",
#                   "emissions_2016"
#                   #"c_emissions_perc",
#                   # "emissions_perkid_2011",
#                   # "emissions_perkid_2016",
#                   #"c_emissions_perkid_perc" 
#                   )) |> arrange(desc(prev_LIMAT_2016)) 
```

```{r include=FALSE}
SAvail_im_COMM_changes |> st_drop_geometry() |> 
  dplyr::select(c("COMMUNITY",
                  "prev_LIMAT_2016", #prevalence of LIM after-tax for households with children under 18
                  "mode", 
                  "kid_proportion_2016", #the proportion of children 5-14 age, sums to 100%
                  "c_kid_proportion_perc", #the % change in proportion between 2016-2011 i.e. (2016-2011)/2011, negative means decrease
                  "V_im_proportion_2016",  #the proportion of V_im, sums to 100%
                  "c_V_im_proportion_perc", #the % change in proportion between 2016-2011 i.e. (2016-2011)/2011, negative means decrease
                  "v_im_2016", #the proportion of V_im (in community) /kid_population (in community) = v_im, sums to 100%
                  "c_v_im_perc", #the % change in proportion between 2016-2011 i.e. (2016-2011)/2011, negative means decrease
                  "total_tt_2016", #summed travel time (kid_pop at each household * travel time of mode, summed for the community). Values sums to 100%
                  "c_total_tt_perc",  #the % change in proportion between 2016-2011 i.e. (2016-2011)/2011, negative means decrease
                  "emissions_perkid_2016", #summed motorized travel time * 0.1 g/min (kid_pop at each household * travel time of mode, summed for the community). Values sums to 100%
                  "c_emissions_perkid_perc"  #the % change in proportion between 2016-2011 i.e. (2016-2011)/2011, negative means decrease
  ))
```

Spatial availability can also be summarized by community. The majority of schools were closed in Hamilton Center, this community represents `r SAvail_im_COMM_changes |> st_drop_geometry() |> filter(COMMUNITY == "HA") |> dplyr::select("kid_proportion_2016") |> sum() |> scales::percent()` of the 5-14 year old population in 2016. Though the population between 2011 and 2016 decreased, the rate of school-seat spatial availability decreased disproportionately more (population by `r SAvail_im_COMM_changes |> st_drop_geometry() |> filter(COMMUNITY == "HA") |> transmute(calc = ((sum(kid_2016)-sum(kid_2011))/sum(kid_2011)) |> scales::percent()) |> first()` and spatial availability by `r `r SAvail_im_COMM_changes |> st_drop_geometry() |> filter(COMMUNITY == "HA") |> transmute(calc = ((sum(V_im_2016)-sum(V_im_2011))/sum(V_im_2011)) |> scales::percent())  |> first()`). This disproportionate decrease is greatest in Hamilton Central than any other community. This may have equity concerns as Hamilton Central has the highest LIM-AT prevalence, with `r SAvail_im_COMM_changes |> st_drop_geometry() |> filter(COMMUNITY == "HA" & mode == "motorized") |> dplyr::select("prev_LIMAT_2016") |> round() |> paste0("%")` of households LIM-AT in 2016, 3.7 times greater than the community with the lowest level of LIM-AT prevalence in the city. Between 2016 and 2011, Hamilton Center had the lowest levels of emissions per student and saw the highest per student increase in 2016. Hamilton Center had the highest non-motorized trip potential, and it was reduced in part by the school closure policies.

Conversely, communities with the lowest LIM-AT benefit the most, notably in gains to motorized spatial availability. Glanbrook, Flamborough and Ancaster have the lowest prevalence of LIM-AT of any Hamilton Community and together account for `r SAvail_im_COMM_changes |> st_drop_geometry() |> filter(COMMUNITY == "FL" | COMMUNITY == "GL" | COMMUNITY == "AN") |> dplyr::select("kid_proportion_2016") |> sum() |> scales::percent()` of the 5-14 year old population in 2016. These communities experienced a growth in student population (or only a small decrease) and proportionally benefited from an increase in motorized spatial availability.

Across the city, those who commute by non-motorized mode see the most dramatic drop in spatial availability across the city. In all communities, spatial availability decreases at greater levels than population. Notably, in Dundas (2nd highest LIM-AT), the only school in the community closed leaving the population in the community with no other option but to use motorized modes. Glanbrook (lowest LIM-AT), which experienced a sharp increase in student population along with non-motorized using population, suffered a lose in non-motorized spatial availability despite seeing the highest increase in motorized spatial availability. Overall, the gain in motorized spatial availability resulted in a further erosion of non-motorized spatial availability. This erosion results in a lose in active transport trips and an increase in emissions city-wide (`r SAvail_im_COMM_changes |> st_drop_geometry() |> transmute(em = (sum(emissions_2011))/(sum(kid_2011)) ) |> sum() |> round(digit=2)`kg $CO_2e$ per home-to-school potential student-trip in 2011 to `r SAvail_im_COMM_changes |> st_drop_geometry() |> transmute(em = (sum(emissions_2016))/(sum(kid_2016)) ) |> sum() |> round(digit=2)` in 2016).

<!-- Is this worth further developing or perhaps just remove, what do you think? I think further. [AP]: tables to show availability by mode in 2011 and 2016 with changes, as well as lost minutes of active travel time and emissions. Check Elise's equity in accessibility to bike share paper, see table 1 there. -->
Study limitations:

- potential students per households, potential trips to all schools
- estimated travel times, direct car or walk trip (shortest path). Not necessarily car, could be transit or school bus. Walking shortest-paths are also optimistic for children's travel: could be unsafe and not taken in reality and/or alternative more pleasant paths are taken.
- estimated emissions. GHG emissions per minute of driving varies extensively depending on operating conditions and vehicle characteristics [@soukhovOccupancyGHGEmissions2022].

<!-- Add this additional paper: it discusses the potential impact of mode-switching on GHG emissions reductions [@pantelakiImpactHomeschoolCommuting2024].

Also, access to schools has broader implications; it has a relationship to residential and job choices [@pietrabissaSchoolAccessCity2023] and housing prices [@merrallWhatSchoolWorth2023] ? Shall I further elaborate in the discussion?-->

# Conclusion

In this paper, we estimate the change in multi-modal spatial availability as a result of school consolidations/closures between the years of 2011 and 2016 in an empirical case study of Hamilton, a mid-size city (\~550,000 pop) in Ontario, Canada. The spatial availability and spatial availability per student are visualized on a DA level; spatial availability for the city declined overall. This decline is then discussed by measuring potential walk minutes and GHG emissions (produced by additional car minutes that result since walk minutes are removed). Overall, the closure of `r ((ALL_SCHOOLS_Elem %>% filter(Year != "2016" & Level == "Elementary") %>% pull(SchoolID) %>% length() - ALL_SCHOOLS_Elem %>% filter(Year != "2011" & Level == "Elementary") %>% pull(SchoolID) %>% length())/ALL_SCHOOLS_Elem %>% filter(Year != "2016" & Level == "Elementary") %>% pull(SchoolID) %>% length()) %>% scales::percent()` of elementary schools in Hamilton resulted in: a `r (ALL_SCHOOLS_Elem %>% st_drop_geometry() %>% filter( (Year == "2011" | Year == "2011 and 2016") & Level == "Elementary") %>% dplyr::select(OTGC2011) %>% sum() - ALL_SCHOOLS_Elem %>% st_drop_geometry() %>% filter( (Year == "2016" | Year == "2011 and 2016") & Level == "Elementary") %>% dplyr::select(OTGC2016) %>% sum()) %>% prettyNum(big.mark=",", digits=1)` decline of school-seat capacity, a decline of `r SAvail_im_COMM_changes |> st_drop_geometry() |> filter(mode == "non-motorized") |> transmute(total_tt = total_tt_2016-total_tt_2011) |> sum() %>% abs() %>% prettyNum(big.mark=",", digits=1)` *potential* home-to-school walk-minutes (per one-way trips), an increase of `r SAvail_im_COMM_changes |> st_drop_geometry() |> filter(mode == "motorized") |> transmute(emissions = emissions_2016-emissions_2011) |> sum() %>% abs() %>% prettyNum(big.mark=",", digits=1)` kg CO2e as a result of increased car-minutes as a result of lost *potential* home-to-school walk-minutes (per one-way trips).

By quantifying the spatial availability, environmental, and active-travel implications of school closure/consolidation policies in Hamilton, we offer a methodology for spatial policy analysis scenario. This methodology can be used by researchers and decision-makers to plan and evaluate equitable and sustainable urban planning policies from a spatial perspective. We urge policy makers to view 'spatial availability' of public resources from a per-capita perspective, plan capacity and location of service provision with the cost of travel (and associated implications such as GHG emissions, active transportation mode, safety of active travel modes, etc.) that sufficiently serves the target population. In this paper we suggest and assume a benchmark of 1.0 school-seats per student which accounts for sufficiency. Spatial availability can be used to obtain this benchmark since the total number of opportunities (i.e., school capacity) is preserved in the region of analysis unlike unconstrained forms of accessibility measurement e.g., Hansen-type measure [@hansenHowAccessibilityShapes1959]. As such, the assigned spatial availability can be divided by population at each zone to yield an interpretable per capita measurement.

Like many other localities, Hamilton is not immune to top-down financial efficient policy decisions to determine community infrastructure is underutilized and it be closed. From the perspective of the provincial government of Ontario, the school closures which occurred in Hamilton resulted in operational savings-per-student but we demonstrate how this resulted in both families and students paying a price through spatial availability, lost potential walk minutes, and increased GHG emissions. In the case of Hamilton Central, where households have the highest prevalence of LIM-AT in the city, those households suffer the most disproportionate decrease in motorized spatial availability and increase in percentage increase in emissions per capacity. As public schools should serve all students; the determination of spatial availability and spatial equity of impact, is critical to better inform decision-makers when evaluating school consolidation policies.

Ultimately, this case study furnishes evidence that consolidation of schools was particularly ill-timed: with the advent of the COVID-19 pandemic, the education system lacked the resiliency to accommodate reduced classroom capacities and left parents who once lived within active transportation distance to schools relying on motorized transportation for their children. These top-down austerity measures left the Hamilton local school system more vulnerable to the impacts of COVID-19 by undervaluing the societal role of schools in neighbourhoods. By identifying the impact of school consolidation policies in Ontario we anticipate that researchers and decision-makers will have better information to center the wellbeing of all residents, including students, and the planet in urban planning policies.

# References

::: {#refs}
:::

# Appendix

\input{TabA1-OTGC-Reg.tex} 
<!--made in the hamONdests repo-->